---
title: "Benchmark tools using pooled database - alpha-beta pairing - CD8 cell subset"
author: "Charline Jouannet"
output: github_document
  html_preview: true
---

```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(echo=TRUE, warning=F, message=F)
```


```{r Library}
library(RCy3)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(ggpubr)
library(viridis)
library(vegan)
library(pheatmap)
library(RColorBrewer)
library(ComplexHeatmap)
```

```{r initiation}
list_method <- c("Hamming_Distance", "Levenshtein_Distance", "TCRMatch", "iSMART","GIANA", "clusTCR","GLIPH2", "DeepTCR", "TCRdist3")
color_method <- c("darkorange", "cyan3", "darkorchid4", "grey", "brown1", "darkgoldenrod1", "pink", "chartreuse3", "black")
```

```{r Database and output Processing}
#process the pooled database
source("../Analysis/database_processing.R")
#process output files from methods
source("../Analysis/output_processing.R")
```

# Creation of network 

```{r network}
#create network
source("../Functions/network_functions.R") 

## Hamming Distance ##
#color by chain type
CDR3_clustered_HD <- data.frame(CDR3=unique(edge_list_HD$input_sequence))
vertices_HD <- merge(CDR3_clustered_HD, liste_sequence_TRA_TRB_chain, by= "CDR3")
#cytoscape
vertices_HD <- merge(vertices_HD, output_HD, by = "CDR3")
graph <- graph_from_data_frame(edge_list_HD, directed = FALSE, vertices = vertices_HD)
#createNetworkFromIgraph(graph)


## Levenshtein Distance ##
#color by chain type
CDR3_clustered_LD <- data.frame(CDR3=unique(edge_list_LD$input_sequence))
vertices_LD <- merge(CDR3_clustered_LD, liste_sequence_TRA_TRB_chain, by= "CDR3")
#cytoscape
vertices_LD <- merge(vertices_LD, output_LD, by = "CDR3")
graph <- graph_from_data_frame(edge_list_LD, directed = FALSE, vertices = vertices_LD)
#createNetworkFromIgraph(graph)


## TCRMatch ##
#color by chain type
CDR3_clustered_tcrmatch <- data.frame(CDR3=unique(edge_list_tcrm$input_sequence))
vertices_tcrmatch <- merge(CDR3_clustered_tcrmatch, liste_sequence_TRA_TRB_chain, by= "CDR3", all.x= T)
#cytoscape
vertices_tcrmatch <- merge(vertices_tcrmatch, output_tcrm, by = "CDR3")
graph <- graph_from_data_frame(edge_list_tcrm, directed = FALSE, vertices = vertices_tcrmatch)
#createNetworkFromIgraph(graph)


## iSMART ##
ismart_list <- split(output_ismart, f = output_ismart$clusters)
edge_list_ismart <- make_edges(ismart_list, "CDR3")#create edges list
edge_list_ismart <- edge_list_ismart %>% filter(value == 1)
colnames(edge_list_ismart) <- c("input_sequence", "match_sequence", "score", "cluster")
edge_list_ismart <- edge_list_ismart %>% select(-cluster)
#color by chain type
CDR3_clustered_ismart <- data.frame(CDR3=unique(output_ismart$CDR3))
vertices_ismart <- merge(CDR3_clustered_ismart, liste_sequence_TRA_TRB_chain, by= "CDR3")
#cytoscape
vertices_ismart <- merge(vertices_ismart , output_ismart, by = "CDR3")
graph <- graph_from_data_frame(edge_list_ismart, directed = FALSE, vertices = vertices_ismart)
#createNetworkFromIgraph(graph)


## GIANA ##
giana_list <- split(output_giana, f = output_giana$clusters)
edge_list_giana <- make_edges(giana_list, "CDR3")
edge_list_giana <- edge_list_giana %>% filter(value == 1)
#color by chain type
CDR3_clustered_giana <- data.frame(CDR3=unique(output_giana$CDR3))
vertices_giana <- merge(CDR3_clustered_giana, liste_sequence_TRA_TRB_chain, by= "CDR3")
#cytoscape
vertices_giana <- merge(vertices_giana, output_giana, by = "CDR3")
graph <- graph_from_data_frame(edge_list_giana, directed = FALSE, vertices = vertices_giana)
#createNetworkFromIgraph(graph)


## clusTCR ##
from_edge_clustcr <- data.frame(CDR3b_CDR3a = unique(edge_list_clustcr$From))
to_edge_clustcr <- data.frame(CDR3b_CDR3a = unique(edge_list_clustcr$To))#combine the two "from" and "to" lists from edge_list of clustcr output
CDR3_clustered_clustcr <- rbind(from_edge_clustcr, to_edge_clustcr) %>% unique() %>% mutate(color_type = "black")
vertices_clustcr <- merge(CDR3_clustered_clustcr, unique_pairing ,by="CDR3b_CDR3a", all.y = T)
vertices_clustcr$color_type[is.na(vertices_clustcr$color_type)] <- "grey"
#cytoscape
vertices_clustcr <- merge(vertices_clustcr, output_clustcr_pairing, by = "CDR3b_CDR3a")
graph <- graph_from_data_frame(edge_list_clustcr, directed = FALSE, vertices = vertices_clustcr)
#createNetworkFromIgraph(graph)


## GLIPH2 ##
##graph## node = pairing alpha-beta
gliph_pairing <- output_gliph %>% select(c(clusters, CDR3b, CDR3a, size, type_cluster))
gliph_pairing_list <- split(gliph_pairing, f = gliph_pairing$clusters)

#color edges based on type of cluster (global/motif)
edge_list_gliph  <- make_edges_gliph2(gliph_pairing_list)
edge_list_gliph  <- edge_list_gliph  %>% filter(value == 1)
# 1.graph annotation
#color by cluster type
CDR3_clustered_gliph <- data.frame(CDR3b_CDR3a=output_gliph$CDR3b_CDR3a, cluster = output_gliph$clusters, type_cluster = output_gliph$type_cluster, size = output_gliph$size)
vertices_gliph_merge <- CDR3_clustered_gliph %>% 
  mutate(color_type = case_when(
    type_cluster == "motif" ~ "blue",
    type_cluster == "global" ~ "grey"))
vertices_gliph_merge$CDR3_cluster <- paste(vertices_gliph_merge$CDR3b_CDR3a, vertices_gliph_merge$cluster, sep = "_")
#color by specificity : highlight one specificity
vertices_gliph_merge <- merge(vertices_gliph_merge, data_human[,c("CDR3b_CDR3a", "Epitope")], by = "CDR3b_CDR3a", all.x = T) %>% unique()
vertices_gliph_list <- split(vertices_gliph_merge, f = vertices_gliph_merge$CDR3_cluster)
vertices_gliph <- color_specificity(vertices_gliph_list, "NLVPMVATV", "red")
vertices_gliph <- vertices_gliph%>% select(-CDR3b_CDR3a) %>% relocate(CDR3_cluster, .before = cluster) %>% unique()
# 2.graph
#cytoscape
vertices_gliph <- vertices_gliph %>% mutate(CDR3b_CDR3a = CDR3_cluster) %>% separate(CDR3b_CDR3a, c('CDR3b_CDR3a', 'clusters_2')) %>% select(-clusters_2)
graph <- graph_from_data_frame(edge_list_gliph, directed = FALSE, vertices = vertices_gliph)
#createNetworkFromIgraph(graph)


## DeepTCR ##
deeptcr_pairing <- output_deeptcr %>% select(c(clusters, CDR3b, CDR3a, size))
deeptcr_pairing$CDR3b_CDR3a <- paste(deeptcr_pairing$CDR3b, deeptcr_pairing$CDR3a, sep = "")
deeptcr_pairing <- deeptcr_pairing %>% select(-c(CDR3b, CDR3a)) %>% unique()
deeptcr_pairing_list <- split(deeptcr_pairing, f = deeptcr_pairing$clusters) 
edge_list_deeptcr <- make_edges(deeptcr_pairing_list, "CDR3b_CDR3a")
edge_list_deeptcr  <- dplyr::rename(edge_list_deeptcr , "CDR3b_CDR3a"="get(column)")
#color by specificity: highlight one specificity
CDR3_clustered_deeptcr <- data.frame(CDR3b_CDR3a = unique(edge_list_deeptcr$CDR3b_CDR3a))
vertices_deeptcr_merge <- merge(CDR3_clustered_deeptcr, data_human[,c("CDR3b_CDR3a", "Epitope")], by = "CDR3b_CDR3a", all.x = T) %>% unique()
vertices_deeptcr_list <- split(vertices_deeptcr_merge, f = vertices_deeptcr_merge$CDR3b_CDR3a)
vertices_deeptcr <- color_specificity(vertices_deeptcr_list, "NLVPMVATV", "red")
#cytoscape
vertices_deeptcr <- merge(vertices_deeptcr, output_deeptcr[c("CDR3b_CDR3a", "clusters", "size")], by = "CDR3b_CDR3a")
graph <- graph_from_data_frame(edge_list_deeptcr, directed = FALSE, vertices = vertices_deeptcr)
#createNetworkFromIgraph(graph)


## TCRdist3 ##
CDR3_clustered_tcrdist <- data.frame(clone = unique(output_tcrdist$clone))
#create edges from hierarchical clustering output using clones
tcrdist_clone <- output_tcrdist %>% select(c(clusters, clone, size)) %>% unique()
tcrdist_clone_list <- split(tcrdist_clone, f = tcrdist_clone$clusters)
edge_list_tcrdist_clone <- make_edges(tcrdist_clone_list, "clone")
edge_list_tcrdist_clone  <- dplyr::rename(edge_list_tcrdist_clone ,"clone"="get(column)") %>% filter(value !=0)
#color by specificity: highlight one specificity
data_human$clone <- paste(data_human$V_alpha, "*01_", data_human$CDR3_alpha, "_", data_human$J_alpha, "_", data_human$V_beta, "*01_", data_human$CDR3_beta, "_", data_human$J_beta, sep = "")
vertices_tcrdist_merge <- merge(CDR3_clustered_tcrdist, data_human[,c("clone", "Epitope")], by.x = "clone", all.x = T) %>% unique()
vertices_tcrdist_list <- split(vertices_tcrdist_merge, f = vertices_tcrdist_merge$clone)
vertices_tcrdist <- color_specificity(vertices_tcrdist_list, "NLVPMVATV", "red")
#cytoscape
vertices_tcrdist <- merge(vertices_tcrdist, output_tcrdist[c("clone", "clusters", "size", "CDR3b_CDR3a")], by = "clone")
graph <- graph_from_data_frame(edge_list_tcrdist_clone, directed = FALSE, vertices = vertices_tcrdist)
#createNetworkFromIgraph(graph)
```


# Size distribution of output clusters

```{r Size distribution, fig.dim=c(12,10)}
cluster_HD <- output_HD %>% select(clusters, size) %>% mutate(method = "Hamming_Distance") %>% unique()
cluster_LD <- output_LD %>% select(clusters, size) %>% mutate(method = "Levenshtein_Distance") %>% unique()
cluster_tcrm <- output_tcrm %>% select(clusters, size) %>% mutate(method = "TCRMatch") %>% unique()
cluster_ismart <- output_ismart %>% select(clusters, size) %>% mutate(method = "iSMART") %>% unique()
cluster_giana <- output_giana %>% select(clusters, size) %>% mutate(method = "GIANA") %>% unique()
cluster_clustcr <- output_clustcr_pairing %>% select(clusters, size) %>% mutate(method = "clusTCR") %>% unique()
cluster_gliph <- output_gliph %>% select(clusters, size) %>% mutate(method = "GLIPH2") %>% unique()
cluster_deeptcr <- output_deeptcr %>% select(clusters, size) %>% mutate(method = "DeepTCR") %>% unique()
cluster_tcrdist <- output_tcrdist %>% select(clusters, size) %>% mutate(method = "TCRdist3") %>% unique()

cluster_all <- list("Hamming_Distance"= cluster_HD, "Levenshtein_Distance"=cluster_LD, "TCRMatch"=cluster_tcrm, "iSMART"=cluster_ismart, "GIANA"=cluster_giana, "clusTCR"=cluster_clustcr, "GLIPH2"=cluster_gliph, "DeepTCR"=cluster_deeptcr, "TCRdist3"=cluster_tcrdist)
color_method <- list("Hamming_Distance"= "orange", "Levenshtein_Distance"= "cyan3", "TCRMatch"= "darkorchid4", "iSMART"= "grey", "GIANA"= "brown1", "clusTCR"= "darkgoldenrod1", "GLIPH2"= "pink", "DeepTCR"= "chartreuse3", "TCRdist3"= "black")

list_plot <- list()
for(method in names(cluster_all)){
  #print(method)
  list_plot[[method]] <- plot_distribution_size(cluster_all[[method]], color_method[[method]], method)
}

#pdf("../Plot/clusterSizeDistribution.pdf", height = 10, width = 10)
gridExtra::grid.arrange(list_plot[["DeepTCR"]], list_plot[["TCRdist3"]], list_plot[["GIANA"]], list_plot[["Hamming_Distance"]], list_plot[["Levenshtein_Distance"]], list_plot[["iSMART"]], list_plot[["GLIPH2"]], list_plot[["TCRMatch"]], list_plot[["clusTCR"]])
#dev.off()

```

# Metrics: retention, purity, perc_clusters, perc_CDR3

```{r}
#import purity functions
source("../Functions/purity_functions.R")
```

```{r purity, retention}
#retention
retention_HD <- round(n_distinct(output_HD$CDR3)/8395,2)#divided by the number of unique sequences
retention_LD <- round(n_distinct(output_LD$CDR3)/8395,2)
retention_tcrm <- round(n_distinct(output_tcrm$CDR3)/8395,2)
retention_ismart <- round(n_distinct(output_ismart$CDR3)/8395,2)
retention_giana <- round(n_distinct(output_giana$CDR3)/8395,2)

retention_gliph <- round(n_distinct(output_gliph[c("CDR3b","CDR3a")])/4779,2)#divided by the number of unique pairs
retention_clustcr <- round(n_distinct(output_clustcr_pairing$CDR3b_CDR3a)/4779,2)
retention_deeptcr <- round(n_distinct(output_deeptcr[c("CDR3b","CDR3a")])/4779,2)
retention_tcrdist <- round(n_distinct(output_tcrdist[c("CDR3b","CDR3a")])/4779,2)

#purity
th <- 1#threshold for cluster size - equal to 1, 3, 5 or 10

purity_HD <- round(purity_function(output_HD, th),2)
purity_LD <- round(purity_function(output_LD, th),2)
purity_tcrm <- round(purity_function(output_tcrm, th),2)
purity_ismart <- round(purity_function(output_ismart, th),2)
purity_giana <- round(purity_function(output_giana, th),2)

purity_clustcr <- round(purity_function(output_clustcr_pairing, th, "CDR3b_CDR3a"),2)
purity_gliph <- round(purity_function(output_gliph, th, "CDR3b_CDR3a"),2)
purity_deeptcr <- round(purity_function(output_deeptcr, th, "CDR3b_CDR3a"),2)
purity_tcrdist <- round(purity_function(output_tcrdist, th, "CDR3b_CDR3a"),2)

data_purity <- data.frame(method = c(list_method),
                          threshold = rep(1, 9),
                          purity = c(purity_HD, purity_LD, purity_tcrm, purity_ismart, purity_giana, purity_clustcr, purity_gliph, purity_deeptcr, purity_tcrdist))


#retention & purity
df_metric <- data.frame(methods   = list_method,
                        retention = c(retention_HD, retention_LD, retention_tcrm, retention_ismart, retention_giana,  retention_clustcr,retention_gliph, retention_deeptcr, retention_tcrdist),
                        purity = c(purity_HD, purity_LD, purity_tcrm, purity_ismart, purity_giana, purity_clustcr, purity_gliph, purity_deeptcr, purity_tcrdist))

df_metric <-  df_metric %>% melt()

df_metric$methods <- factor(df_metric$methods, level = list_method)
df_metric$variable<- factor(df_metric$variable, level = c("retention", "purity"))

#barplot
# ggplot(df_metric, aes(x = variable, y = value, fill = methods)) +
#   geom_bar(stat = "identity", position = "dodge", color = "white") +
#   theme_classic() +
#   scale_fill_manual(values = c("darkorange", "cyan3", "darkorchid4", "grey", "brown1", "pink", "darkgoldenrod1", "chartreuse3", "black"))+
#   geom_text(aes(label = round(value, 2)), vjust = -0.5,
#             position = position_dodge(0.9)) +
#   scale_y_continuous(expand = c(0.0,0.01), limits = c(0,1.1))+
#   theme(legend.position ="top")+
#   labs(title = "Retention & Purity")

#bubble plot
# ggplot(df_metric, aes(x=variable, y = methods, color = methods, size = value)) +
#   geom_point(alpha=0.5) +
#   scale_size(range = c(2, 12))+
#   theme_classic() +
#   scale_color_manual(values = color_method)+
#   geom_text(data=df_metric, aes(label=value), size=4, color = "black")
```

# Correlation between purity and retention

```{r correlation between metrics}
#correlation purity/retention
corr_purity_retention <- data.frame(methods   = list_method,
                        retention = c(retention_HD, retention_LD, retention_tcrm, retention_ismart, retention_giana, retention_clustcr, retention_gliph, retention_deeptcr, retention_tcrdist),
                        purity = c(purity_HD, purity_LD, purity_tcrm, purity_ismart, purity_giana, purity_clustcr, purity_gliph, purity_deeptcr, purity_tcrdist))

#pdf("../Plot/correlationPurityRetention.pdf")
ggscatter(corr_purity_retention, x = "retention", y = "purity", 
          add = "reg.line", conf.int = TRUE, label = "methods", repel = TRUE, cor.coeff.args = list(label.x = 0.1 , label.y = 0.1 ),
          cor.coef = TRUE, cor.method = "pearson",cor.coef.size = 6,
          xlab = "Retention", ylab = "Purity", font.label = c(15, "plain"))+
  theme(legend.position = "right")+
 font("xlab", size = 15)+
 font("ylab", size = 15)+
 font("xy.text", size = 12)
#dev.off()

#correlation between purity and cluster size
# library(smplot2)
# for(method_1 in list_method){
#   subset <- purity_all_df %>% filter(method == method_1)
#   print(ggplot(data = subset, mapping = aes(x = size, y = purity)) +
#   geom_point(shape = 21, fill = color_method[[method_1]], color = 'white', size = 3) +
#   sm_statCorr()+ggtitle(method_1))#pearson correlation by default
# }
# 
# 
# ggscatter(purity_all_df, x = "size", y = "purity", 
#           add = "reg.line", conf.int = TRUE, 
#           cor.coef = TRUE, cor.method = "pearson",
#           xlab = "Cluster size", ylab = "Purity",
#           color = "method", palette = c("darkorange", "cyan3", "darkorchid4", "grey", "brown1",  "darkgoldenrod1", "pink", "chartreuse3", "black"),
#            facet.by = "method")
```

# Calculation of purity for each cluster to determine the percentage with purity >0.9

```{r purity by cluster, fig.dim=c(8,10)}
#1) Purity for each cluster
purity_HD_df <- all_purity_onechain(output_HD, th) %>% mutate(method = "Hamming_Distance") 
purity_LD_df <- all_purity_onechain(output_LD, th) %>% mutate(method = "Levenshtein_Distance")
purity_tcrm_df <- all_purity_onechain(output_tcrm, th) %>% mutate(method = "TCRMatch")
purity_ismart_df <- all_purity_onechain(output_ismart, th) %>% mutate(method = "iSMART") 
purity_giana_df <- all_purity_onechain(output_giana, th) %>% mutate(method = "GIANA")

purity_clustcr_df  <- all_purity_pairing(output_clustcr_pairing, "CDR3b_CDR3a") %>% mutate(method = "clusTCR")
purity_gliph_df  <- all_purity_pairing(output_gliph, "CDR3b_CDR3a") %>% mutate(method = "GLIPH2") 
purity_deeptcr_df  <- all_purity_pairing(output_deeptcr, "CDR3b_CDR3a") %>% mutate(method = "DeepTCR")
purity_tcrdist_df <- all_purity_pairing(output_tcrdist, "CDR3b_CDR3a") %>% mutate(method = "TCRdist3")

# purity_all_df <- rbind(purity_HD_df, purity_LD_df, purity_tcrm_df, purity_ismart_df ,purity_giana_df,
#                         purity_clustcr_df, purity_gliph_df, purity_deeptcr_df, purity_tcrdist_df)
# purity_all_df$method <- factor(purity_all_df$method, level = list_method)
# 
#boxplot
# ggplot(purity_all_df, aes(x=method, y=purity)) + 
#   geom_boxplot(trim=FALSE, aes(fill = method), alpha = 0.7)+
#   scale_fill_manual(values = c("darkorange", "cyan3", "darkorchid4", "grey", "brown1", "pink", "darkgoldenrod1", "chartreuse3", "black"))+
#   #geom_boxplot(width=0.1)+
#   stat_summary(fun.data=mean_sdl, 
#                  geom="pointrange", color="black")+
#  
#   geom_jitter(shape=16, position=position_jitter(0.2), aes(color = size))+
#   scale_color_continuous(type = "viridis", trans = "reverse")+
#   theme_classic()+
#   ggtitle("Local Purity")

#2) Percentage of clusters (with purity >0.9) and percentage of CDR3 sequences/pairs in these clusters
purity_gliph_df  <- merge(purity_gliph_df, output_gliph[,c("clusters", "CDR3b_CDR3a")], by = "clusters")
purity_tcrdist_df  <- merge(purity_tcrdist_df, output_tcrdist[,c("clusters", "CDR3b_CDR3a")], by = "clusters")

purity_clusters <- data.frame(methods = list_method,
                             perc_clusters = c(n_distinct(filter(purity_HD_df, purity >= 0.9)$clusters)/n_distinct(purity_HD_df$clusters),
                                               n_distinct(filter(purity_LD_df, purity >= 0.9)$clusters)/n_distinct(purity_LD_df$clusters),
                                               n_distinct(filter(purity_tcrm_df, purity >=0.9)$clusters)/n_distinct(purity_tcrm_df$clusters),
                                               n_distinct(filter(purity_ismart_df, purity >= 0.9)$clusters)/n_distinct(purity_ismart_df$clusters),
                                               n_distinct(filter(purity_giana_df, purity >= 0.9)$clusters)/n_distinct(purity_giana_df$clusters),
                                               n_distinct(filter(purity_clustcr_df, purity >= 0.9)$clusters)/n_distinct(purity_clustcr_df$clusters),
                                               n_distinct(filter(purity_gliph_df, purity >= 0.9)$clusters)/n_distinct(purity_gliph_df$clusters),
                                               n_distinct(filter(purity_deeptcr_df, purity >= 0.9)$clusters)/n_distinct(purity_deeptcr_df$clusters),
                                               n_distinct(filter(purity_tcrdist_df, purity >= 0.9)$clusters)/n_distinct(purity_tcrdist_df$clusters)),
                             perc_CDR3s =  c(sum(filter(purity_HD_df, purity >= 0.9)$size)/sum(purity_HD_df$size),
                                             sum(filter(purity_LD_df, purity >= 0.9)$size)/sum(purity_LD_df$size),
                                             sum(filter(purity_tcrm_df, purity >= 0.9)$size)/sum(purity_tcrm_df$size),
                                             sum(filter(purity_ismart_df, purity >= 0.9)$size)/sum(purity_ismart_df$size),
                                             sum(filter(purity_giana_df, purity >= 0.9)$size)/sum(purity_giana_df$size),
                                             sum(filter(purity_clustcr_df, purity >= 0.9)$size)/sum(purity_clustcr_df$size),
                                             n_distinct(filter(purity_gliph_df, purity >= 0.9)$CDR3b_CDR3a)/n_distinct(purity_gliph_df$CDR3b_CDR3a),
                                             sum(filter(purity_deeptcr_df, purity >= 0.9)$size)/sum(purity_deeptcr_df$size),
                                             n_distinct(filter(purity_tcrdist_df, purity >= 0.9)$CDR3b_CDR3a)/n_distinct(purity_tcrdist_df$CDR3b_CDR3a)))
                             

purity_clusters <- melt(purity_clusters)

purity_clusters$methods <- factor(purity_clusters$methods, level = list_method)

#barplot 
# ggplot(purity_clusters, aes(x = variable, y = value, fill = methods)) +
#   geom_bar(stat = "identity", position = "dodge", color = "white") +
#   theme_bw() +
#   scale_fill_manual(values = c("darkorange", "cyan3", "darkorchid4", "grey", "brown1", "darkgoldenrod1", "pink", "chartreuse3", "black"))+
#   geom_text(aes(label = round(value, 2)), vjust = -0.5,
#             position = position_dodge(0.9)) +
#   #scale_y_continuous(expand = c(0,100), limits = c(0,100))+
#   #scale_y_continuous(expand = c(0.0,0.01), limits = c(0,1.1))+
#   theme(legend.position ="top")+
#   labs(title = "Purity (>=0.9)")+
#   ylab("Fraction")


#Bubbleplot of all metrics: retention, purity, perc_clusters (purity >0.9) and perc_CDR3 in these clusters
purity_clusters$value <- round(purity_clusters$value,2)
df_all_metric <- rbind(df_metric, purity_clusters)

# df_all_metric$methods <- factor(df_all_metric$methods, level = rev(list_method))

#pdf("../Plot/AllMetricsTh_1.pdf", width = 10)
# ggplot(df_all_metric, aes(x=variable, y = methods, color = value, size = value)) +
#   geom_point(alpha=0.7) +
#   scale_size(range = c(2, 12))+
#   theme_bw() +
#   scale_color_viridis(discrete = F, direction = -1)+
#   geom_text(data=df_all_metric, aes(label=value), size=4, color = "black",vjust=-2)+
#   theme(legend.position ="right",
#         plot.title = element_text(hjust=0.5),
#         legend.text=element_text(size=15),
#         legend.title=element_text(size=15),
#         text=element_text(size=15),
#         axis.text=element_text(size=15))+
#   ylab("")+
#   xlab("")+
#   guides(color=guide_legend(), size = guide_legend())+
#   ggtitle(paste0("cluster_size >", th))
#dev.off()

df_all_metric$methods <- factor(df_all_metric$methods, level = list_method)
#barplot
#pdf("../Plot/AllMetricsTh_1.pdf", width =15, height = 6)
ggplot(df_all_metric, aes(x=variable, y = value, fill = methods)) +
  geom_bar(stat = "identity", position = "dodge", color = "white", alpha=0.7) +
  theme_bw() +
  scale_fill_manual(values = c("darkorange", "cyan3", "darkorchid4", "grey", "brown1", "darkgoldenrod1", "pink", "chartreuse3", "black"))+
  geom_text(aes(label = round(value, 2)), vjust = -0.5,
             position = position_dodge(0.9)) +
  theme(legend.position ="bottom",
        plot.title = element_text(hjust=0.5),
        legend.text=element_text(size=15),
        legend.title=element_text(size=15),
        text=element_text(size=15),
        axis.text=element_text(size=15))+
  ylab("")+
  xlab("")+
  guides(color=guide_legend(), size = guide_legend())+
  ggtitle(paste0("cluster_size >", th))
#dev.off()
```

# purity in function of the chain type

```{r purity for chain}
#check if clusters are homogeneous according chain type
purity_HD <- purity_chain_function(output_HD)
purity_LD <- purity_chain_function(output_LD)
purity_tcrm <- purity_chain_function(output_tcrm)
purity_ismart <- purity_chain_function(output_ismart)
purity_giana <- purity_chain_function(output_giana)

df_purity_chain <- data.frame(methods   = c("Hamming_Distance", "Levenshtein_Distance", "TCRMatch", "iSMART", "GIANA"),
                        purity = c(purity_HD, purity_LD, purity_tcrm, purity_ismart, purity_giana))
df_purity_chain$methods <- factor(df_purity_chain$methods, level = c("Hamming_Distance", "Levenshtein_Distance", "TCRMatch", "iSMART", "GIANA"))

#pdf("../Plot/purityChain.pdf")
ggplot(df_purity_chain, aes(x=methods, y=purity, fill=methods)) +
    #geom_segment(aes(xend=methods, yend=0)) +
    #geom_point(size=4, color=c("darkorange", "cyan3", "darkorchid4", "grey", "brown1")) +
  geom_bar(stat = "identity", position = "dodge", color = "white", alpha=0.7) +
    theme_bw() +
    xlab("")+
  scale_fill_manual(values = c("darkorange", "cyan3", "darkorchid4", "grey", "brown1"))+
  geom_text(aes(label = round(purity, 2)), vjust=1.6,
             position = position_dodge(0.9)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1),
        legend.text=element_text(size=15),
        legend.title=element_text(size=15),
        text=element_text(size=15),
        axis.text=element_text(size=15), 
        legend.position="none")+
  ylab("Purity")
#dev.off()
```

# Evolution of metrics when threshold varie (not in paper)

```{r evolution of purity when th varies}
#1. Purity
threshold <- c(1, 3, 5, 10)
data_purity_all <- data.frame()
for(th in threshold){
  #print(th)
  purity_HD <- round(purity_function(output_HD, th),2)
  purity_LD <- round(purity_function(output_LD, th),th)
  purity_tcrm <- round(purity_function(output_tcrm, th),th)
  purity_ismart <- round(purity_function(output_ismart, th),th)
  purity_giana <- round(purity_function(output_giana, th),th)
  
  purity_clustcr <- round(purity_function(output_clustcr_pairing, th, "CDR3b_CDR3a"),th)
  purity_gliph <- round(purity_function(output_gliph, th, "CDR3b_CDR3a"),th)
  purity_deeptcr <- round(purity_function(output_deeptcr, th, "CDR3b_CDR3a"),th)
  purity_tcrdist <- round(purity_function(output_tcrdist, th, "CDR3b_CDR3a"),th)
  
  data_purity <- data.frame(method = c(list_method),
                          threshold = rep(th, 9),
                          purity = c(purity_HD, purity_LD, purity_tcrm, purity_ismart, purity_giana, purity_clustcr, purity_gliph, purity_deeptcr, purity_tcrdist))
  data_purity_all <- rbind(data_purity_all, data_purity)
  
}

data_purity_all$method <- factor(data_purity_all$method, levels = list_method)
data_purity_all$threshold <- as.character(data_purity_all$threshold)
data_purity_all$threshold[data_purity_all$threshold == "1"] <- "all"
data_purity_all$threshold <- factor(data_purity_all$threshold, levels = c("all", "3", "5", "10"))

ggplot(data = data_purity_all, aes(x = threshold, y = purity, color = method, group = method))+
  geom_point()+geom_line()+
  theme_bw()+
  scale_color_manual(values = c("darkorange", "cyan3", "darkorchid4", "grey", "brown1", "darkgoldenrod1", "pink", "chartreuse3", "black"))+
  ylim(0,1)

#2. Percentage of clusters/CDR3s
percentage_cluster_all <- data.frame()
percentage_CDR3s_all <- data.frame()
for(th in threshold){
  #print(th)
  purity_HD_df <- all_purity_onechain(output_HD, th) %>% mutate(method = "Hamming_Distance") %>% mutate(method_type = "alpha_or_beta")
  purity_LD_df <- all_purity_onechain(output_LD, th) %>% mutate(method = "Levenshtein_Distance") %>% mutate(method_type = "alpha_or_beta")
  purity_tcrm_df <- all_purity_onechain(output_tcrm, th) %>% mutate(method = "TCRMatch") %>% mutate(method_type = "alpha_or_beta")
  purity_ismart_df <- all_purity_onechain(output_ismart, th) %>% mutate(method = "iSMART") %>% mutate(method_type = "alpha_or_beta")
  purity_giana_df <- all_purity_onechain(output_giana, th) %>% mutate(method = "GIANA") %>% mutate(method_type = "alpha_or_beta")
  
  purity_clustcr_df  <- all_purity_pairing(output_clustcr_pairing, "CDR3b_CDR3a") %>% mutate(method = "clusTCR") %>% mutate(method_type = "alpha_beta_pairing")
  purity_gliph_df  <- all_purity_pairing(output_gliph, "CDR3b_CDR3a") %>% mutate(method = "GLIPH2") %>% mutate(method_type = "alpha_beta_pairing")
  purity_deeptcr_df  <- all_purity_pairing(output_deeptcr, "CDR3b_CDR3a") %>% mutate(method = "DeepTCR") %>% mutate(method_type = "alpha_beta_pairing")
  purity_tcrdist_df <- all_purity_pairing(output_tcrdist, "CDR3b_CDR3a") %>% mutate(method = "TCRdist") %>% mutate(method_type = "alpha_beta_pairing")
  
  purity_all_df <- rbind(purity_HD_df, purity_LD_df, purity_tcrm_df, purity_ismart_df ,purity_giana_df,
                          purity_clustcr_df, purity_gliph_df, purity_deeptcr_df, purity_tcrdist_df)
  purity_all_df$method <- factor(purity_all_df$method, level = c("Hamming_Distance", "Levenshtein_Distance", "TCRMatch", "iSMART", "GIANA", "GLIPH2", "clusTCR", "DeepTCR", "TCRdist"))
  
  purity_gliph_df  <- merge(purity_gliph_df, output_gliph[,c("clusters", "CDR3b_CDR3a")], by = "clusters")
  purity_tcrdist_df  <- merge(purity_tcrdist_df, output_tcrdist[,c("clusters", "CDR3b_CDR3a")], by = "clusters")
  
  purity_clusters <- data.frame(methods = c("Hamming_Distance", "Levenshtein_Distance", "TCRMatch", "iSMART", "GIANA", "clusTCR","GLIPH2", "DeepTCR", "TCRdist"),
                               perc_clusters = c(n_distinct(filter(purity_HD_df, purity >= 0.9)$clusters)/n_distinct(purity_HD_df$clusters),
                                                 n_distinct(filter(purity_LD_df, purity >= 0.9)$clusters)/n_distinct(purity_LD_df$clusters),
                                                 n_distinct(filter(purity_tcrm_df, purity >=0.9)$clusters)/n_distinct(purity_tcrm_df$clusters),
                                                 n_distinct(filter(purity_ismart_df, purity >= 0.9)$clusters)/n_distinct(purity_ismart_df$clusters),
                                                 n_distinct(filter(purity_giana_df, purity >= 0.9)$clusters)/n_distinct(purity_giana_df$clusters),
                                                 n_distinct(filter(purity_clustcr_df, purity >= 0.9)$clusters)/n_distinct(purity_clustcr_df$clusters),
                                                 n_distinct(filter(purity_gliph_df, purity >= 0.9)$clusters)/n_distinct(purity_gliph_df$clusters),
                                                 n_distinct(filter(purity_deeptcr_df, purity >= 0.9)$clusters)/n_distinct(purity_deeptcr_df$clusters),
                                                 n_distinct(filter(purity_tcrdist_df, purity >= 0.9)$clusters)/n_distinct(purity_tcrdist_df$clusters)),
                               perc_CDR3s =  c(sum(filter(purity_HD_df, purity >= 0.9)$size)/sum(purity_HD_df$size),
                                               sum(filter(purity_LD_df, purity >= 0.9)$size)/sum(purity_LD_df$size),
                                               sum(filter(purity_tcrm_df, purity >= 0.9)$size)/sum(purity_tcrm_df$size),
                                               sum(filter(purity_ismart_df, purity >= 0.9)$size)/sum(purity_ismart_df$size),
                                               sum(filter(purity_giana_df, purity >= 0.9)$size)/sum(purity_giana_df$size),
                                               sum(filter(purity_clustcr_df, purity >= 0.9)$size)/sum(purity_clustcr_df$size),
                                               n_distinct(filter(purity_gliph_df, purity >= 0.9)$CDR3b_CDR3a)/n_distinct(purity_gliph_df$CDR3b_CDR3a),
                                               sum(filter(purity_deeptcr_df, purity >= 0.9)$size)/sum(purity_deeptcr_df$size),
                                               n_distinct(filter(purity_tcrdist_df, purity >= 0.9)$CDR3b_CDR3a)/n_distinct(purity_tcrdist_df$CDR3b_CDR3a)))
                               
  
  
  purity_clusters <- purity_clusters %>% mutate(threshold = rep(th,9))
  percentage_cluster <- purity_clusters[,c("methods", "threshold", "perc_clusters")]
  percentage_CDR3s <- purity_clusters[,c("methods", "threshold", "perc_CDR3s")]
  
  percentage_cluster_all <- rbind(percentage_cluster_all, percentage_cluster)
  percentage_CDR3s_all <- rbind(percentage_CDR3s_all, percentage_CDR3s)
}

percentage_cluster_all$methods <- factor(percentage_cluster_all$methods, levels = list_method)
percentage_cluster_all$threshold <- as.character(percentage_cluster_all$threshold)
percentage_cluster_all$threshold[ percentage_cluster_all$threshold == "1"] <- "all"
percentage_cluster_all$threshold <- factor(percentage_cluster_all$threshold, levels = c("all", "3", "5", "10"))

percentage_CDR3s_all$methods <- factor(percentage_CDR3s_all$methods, levels = list_method)
percentage_CDR3s_all$threshold <- as.character(percentage_CDR3s_all$threshold)
percentage_CDR3s_all$threshold[ percentage_CDR3s_all$threshold == "1"] <- "all"
percentage_CDR3s_all$threshold <- factor(percentage_CDR3s_all$threshold, levels = c("all", "3", "5", "10"))

ggplot(data = percentage_CDR3s_all, aes(x = threshold, y = perc_CDR3s, color = methods, group = methods))+
  geom_point()+geom_line()+
  theme_bw()+
  scale_color_manual(values = c("darkorange", "cyan3", "darkorchid4", "grey", "brown1", "darkgoldenrod1", "pink", "chartreuse3", "black"))+
  ylim(0,1)+
  ylab("Percentage (%)")
```

# Re-Calculation of metrics for alpha and beta independently

```{r metrics separating alpha & beta clusters, fig.dim=c(12,10)}
th <- 1
#retention
retention_HD_alpha <- round(n_distinct(output_HD_alpha$CDR3)/4103,2)#divided by the number of unique alpha sequences
retention_LD_alpha <- round(n_distinct(output_LD_alpha$CDR3)/4103,2)
retention_tcrm_alpha <- round(n_distinct(output_tcrm_alpha$CDR3)/4103,2)
retention_ismart_alpha <- round(n_distinct(output_ismart_alpha$CDR3)/4103,2)
retention_giana_alpha <- round(n_distinct(output_giana_alpha$CDR3)/4103,2)

retention_HD_beta <- round(n_distinct(output_HD_beta$CDR3)/4292,2)#divided by the number of unique beta sequences
retention_LD_beta <- round(n_distinct(output_LD_beta$CDR3)/4292,2)
retention_tcrm_beta <- round(n_distinct(output_tcrm_beta$CDR3)/4292,2)
retention_ismart_beta <- round(n_distinct(output_ismart_beta$CDR3)/4292,2)
retention_giana_beta <- round(n_distinct(output_giana_beta$CDR3)/4292,2)

#purity
purity_HD_alpha <- round(purity_function(output_HD_alpha, th),2)
purity_LD_alpha <- round(purity_function(output_LD_alpha, th),2)
purity_tcrm_alpha <- round(purity_function(output_tcrm_alpha, th),2)
purity_ismart_alpha <- round(purity_function(output_ismart_alpha, th),2)
purity_giana_alpha <- round(purity_function(output_giana_alpha, th),2)

purity_HD_beta <- round(purity_function(output_HD_beta, th),2)
purity_LD_beta <- round(purity_function(output_LD_beta, th),2)
purity_tcrm_beta <- round(purity_function(output_tcrm_beta, th),2)
purity_ismart_beta <- round(purity_function(output_ismart_beta, th),2)
purity_giana_beta <- round(purity_function(output_giana_beta, th),2)

#purity by clusters
purity_HD_df_alpha <- all_purity_onechain(output_HD_alpha, th) %>% mutate(method = "Hamming_Distance") %>% mutate(chain = "alpha")
purity_LD_df_alpha <- all_purity_onechain(output_LD_alpha, th) %>% mutate(method = "Levenshtein_Distance") %>% mutate(chain = "alpha")
purity_tcrm_df_alpha <- all_purity_onechain(output_tcrm_alpha, th) %>% mutate(method = "TCRMatch") %>% mutate(chain = "alpha")
purity_ismart_df_alpha <- all_purity_onechain(output_ismart_alpha, th) %>% mutate(method = "iSMART")%>% mutate(chain = "alpha")
purity_giana_df_alpha <- all_purity_onechain(output_giana_alpha, th) %>% mutate(method = "GIANA") %>% mutate(chain = "alpha")

purity_HD_df_beta <- all_purity_onechain(output_HD_beta, th) %>% mutate(method = "Hamming_Distance") %>% mutate(chain = "beta")
purity_LD_df_beta <- all_purity_onechain(output_LD_beta, th) %>% mutate(method = "Levenshtein_Distance") %>% mutate(chain = "beta")
purity_tcrm_df_beta <- all_purity_onechain(output_tcrm_beta, th) %>% mutate(method = "TCRMatch") %>% mutate(chain = "beta")
purity_ismart_df_beta <- all_purity_onechain(output_ismart_beta, th) %>% mutate(method = "iSMART")%>% mutate(chain = "beta")
purity_giana_df_beta <- all_purity_onechain(output_giana_beta, th) %>% mutate(method = "GIANA") %>% mutate(chain = "beta")

#Percentage of clusters (purity >0.9) and percentage of correctly clustered unique CDR3β sequences
purity_clusters_alpha <- data.frame(methods = c("Hamming_Distance", "Levenshtein_Distance", "TCRMatch", "iSMART", "GIANA"),
                             retention = c(retention_HD_alpha, retention_LD_alpha, retention_tcrm_alpha, retention_ismart_alpha, retention_giana_alpha), 
                             purity= c(purity_HD_alpha, purity_LD_alpha, purity_tcrm_alpha, purity_ismart_alpha, purity_giana_alpha),
                             perc_clusters = c(n_distinct(filter(purity_HD_df_alpha, purity >= 0.9)$clusters)/n_distinct(purity_HD_df_alpha$clusters),
                                               n_distinct(filter(purity_LD_df_alpha, purity >= 0.9)$clusters)/n_distinct(purity_LD_df_alpha$clusters),
                                               n_distinct(filter(purity_tcrm_df_alpha, purity >=0.9)$clusters)/n_distinct(purity_tcrm_df_alpha$clusters),
                                               n_distinct(filter(purity_ismart_df_alpha, purity >= 0.9)$clusters)/n_distinct(purity_ismart_df_alpha$clusters),
                                               n_distinct(filter(purity_giana_df_alpha, purity >= 0.9)$clusters)/n_distinct(purity_giana_df_alpha$clusters)
                                               ),
                             perc_CDR3s =  c(sum(filter(purity_HD_df_alpha, purity >= 0.9)$size)/sum(purity_HD_df_alpha$size),
                                             sum(filter(purity_LD_df_alpha, purity >= 0.9)$size)/sum(purity_LD_df_alpha$size),
                                             sum(filter(purity_tcrm_df_alpha, purity >= 0.9)$size)/sum(purity_tcrm_df_alpha$size),
                                             sum(filter(purity_ismart_df_alpha, purity >= 0.9)$size)/sum(purity_ismart_df_alpha$size),
                                             sum(filter(purity_giana_df_alpha, purity >= 0.9)$size)/sum(purity_giana_df_alpha$size)),
                             chain = rep("alpha", 5)
                                             )

purity_clusters_beta <- data.frame(methods = c("Hamming_Distance", "Levenshtein_Distance", "TCRMatch", "iSMART", "GIANA"),
                             retention = c(retention_HD_beta, retention_LD_beta, retention_tcrm_beta, retention_ismart_beta, retention_giana_beta),
                             purity=c(purity_HD_beta, purity_LD_beta, purity_tcrm_beta, purity_ismart_beta, purity_giana_beta),
                             perc_clusters = c(n_distinct(filter(purity_HD_df_beta, purity >= 0.9)$clusters)/n_distinct(purity_HD_df_beta$clusters),
                                               n_distinct(filter(purity_LD_df_beta, purity >= 0.9)$clusters)/n_distinct(purity_LD_df_beta$clusters),
                                               n_distinct(filter(purity_tcrm_df_beta, purity >=0.9)$clusters)/n_distinct(purity_tcrm_df_beta$clusters),
                                               n_distinct(filter(purity_ismart_df_beta, purity >= 0.9)$clusters)/n_distinct(purity_ismart_df_beta$clusters),
                                               n_distinct(filter(purity_giana_df_beta, purity >= 0.9)$clusters)/n_distinct(purity_giana_df_beta$clusters)
                                               ),
                             perc_CDR3s =  c(sum(filter(purity_HD_df_beta, purity >= 0.9)$size)/sum(purity_HD_df_beta$size),
                                             sum(filter(purity_LD_df_beta, purity >= 0.9)$size)/sum(purity_LD_df_beta$size),
                                             sum(filter(purity_tcrm_df_beta, purity >= 0.9)$size)/sum(purity_tcrm_df_beta$size),
                                             sum(filter(purity_ismart_df_beta, purity >= 0.9)$size)/sum(purity_ismart_df_beta$size),
                                             sum(filter(purity_giana_df_beta, purity >= 0.9)$size)/sum(purity_giana_df_beta$size)),
                             chain = rep("beta", 5)
                                             )
                             


purity_clusters_alpha <- melt(purity_clusters_alpha)
purity_clusters_beta <- melt(purity_clusters_beta)

purity_clusters_chain <- rbind(purity_clusters_alpha, purity_clusters_beta)

list_plot <- list()
for(method in unique(purity_clusters_chain$method)){
  purity_clusters_chain_subset <- purity_clusters_chain %>% filter(methods== method)
  
  list_plot[[method]] <- ggplot(data = purity_clusters_chain_subset, aes(x=variable, y = value, fill = chain))+
    geom_bar(stat="identity", position=position_dodge())+
    theme_bw()+
    ggtitle(method)+
    scale_fill_manual(values= alpha(c("chartreuse3","darkorange2"), 0.7))+
    theme(plot.title = element_text(hjust = 0.5, face = "bold"))+
    ylab("")+xlab("")+ylim(0,1)
  
  legend <- get_legend(list_plot[[method]])
  list_plot[[method]] <- list_plot[[method]]+theme(legend.position="none")

}
#pdf("../Plot/AllMetricsTh_1_byChain.pdf")
grid.arrange(list_plot[["Hamming_Distance"]], list_plot[["Levenshtein_Distance"]], list_plot[["TCRMatch"]], list_plot[["iSMART"]] , list_plot[["GIANA"]],
             legend, ncol=3)
#dev.off()
```

# Number of output clusters for each cluster size threshold

```{r number of clusters for each size threshold}
#Supplementary Table 3
list_output <- list("Hamming_Distance" = output_HD, "Levenshtein_Distance"=output_LD, "TCRMatch"= output_tcrm, "GIANA"= output_giana, "iSMART" = output_ismart, "clusTCR" = output_clustcr_pairing, "GLIPH2" = output_gliph,"DeepTCR" = output_deeptcr, "TCRdist3" = output_tcrdist)
th_list <- c(3,5,10)
df_nb_cluster_all <- data.frame()
for(th in th_list){
  #print(th)
  df_nb_cluster <- data.frame()
  for(method in names(list_output)){
    output <- filter(list_output[[method]], size > th)
    df <- data.frame(methods = method, nb_clusters = n_distinct(output$clusters), threshold =th) 
    df_nb_cluster <- rbind(df_nb_cluster, df)
  }
  df_nb_cluster_all <- rbind(df_nb_cluster_all, df_nb_cluster)
}
```

# Overlap of clustered sequences between methods (Upset plot) (not in paper)

```{r upset plot}
list_method_1 <- c("Hamming_Distance", "Levenshtein_Distance", "TCRMatch", "iSMART", "GIANA")
list_method_2 <- c("clusTCR", "GLIPH2", "DeepTCR", "TCRdist3")
list_method <- c("Hamming_Distance", "Levenshtein_Distance", "TCRMatch", "iSMART", "GIANA", "clusTCR", "GLIPH2", "DeepTCR", "TCRdist3")

#list_method_1
m_alpha <- make_comb_mat(Hamming_Distance = output_HD_alpha$CDR3, Levenshtein_Distance = output_LD_alpha$CDR3, TCRMatch = output_tcrm_alpha$CDR3, iSMART = output_ismart_alpha$CDR3, GIANA = output_giana_alpha$CDR3, mode = "intersect")

m_beta <- make_comb_mat(Hamming_Distance = output_HD_beta$CDR3, Levenshtein_Distance = output_LD_beta$CDR3, TCRMatch = output_tcrm_beta$CDR3, iSMART = output_ismart_beta$CDR3, GIANA = output_giana_beta$CDR3, mode = "intersect")

color_set <- c("darkorange" = "Hamming_Distance", 
               "cyan3" = "Levenshtein_Distance",
               "darkorchid4" = "TCRMatch",
               "grey" = "iSMART",
               "brown3" = "GIANA")

#list_method_2
# m_alpha <- make_comb_mat(clusTCR = output_clustcr_alpha$CDR3, GLIPH2 = output_gliph_alpha$CDR3, DeepTCR = output_deeptcr_alpha$CDR3, TCRdist3 = output_tcrdist_alpha$CDR3, mode = "intersect")
# 
# m_beta <- make_comb_mat(clusTCR = output_clustcr_beta$CDR3, GLIPH2 = output_gliph_beta$CDR3, DeepTCR = output_deeptcr_beta$CDR3, TCRdist3 = output_tcrdist_beta$CDR3, mode = "intersect")
# 
# color_set <- c("darkgoldenrod1" = "clusTCR", 
#                "lightpink" = "GLIPH2",
#                "chartreuse3" = "DeepTCR",
#                "black" = "TCRdist3")
#all
# m_alpha <- make_comb_mat(Hamming_Distance = output_HD_alpha$CDR3, Levenshtein_Distance = output_LD_alpha$CDR3, TCRMatch = output_tcrm_alpha$CDR3, iSMART = output_ismart_alpha$CDR3, GIANA = output_giana_alpha$CDR3, clusTCR = output_clustcr_alpha$CDR3, GLIPH2 = output_gliph_alpha$CDR3, DeepTCR = output_deeptcr_alpha$CDR3, TCRdist3 = output_tcrdist_alpha$CDR3, mode = "intersect")
# 
# m_beta <- make_comb_mat(Hamming_Distance = output_HD_beta$CDR3, Levenshtein_Distance = output_LD_beta$CDR3, TCRMatch = output_tcrm_beta$CDR3, iSMART = output_ismart_beta$CDR3, GIANA = output_giana_beta$CDR3, clusTCR = output_clustcr_beta$CDR3, GLIPH2 = output_gliph_beta$CDR3, DeepTCR = output_deeptcr_beta$CDR3, TCRdist3 = output_tcrdist_beta$CDR3, mode = "intersect")
# 
# color_set <- c("darkorange" = "Hamming_Distance", 
#                "cyan3" = "Levenshtein_Distance",
#                "darkorchid4" = "TCRMatch",
#                "grey" = "iSMART",
#                "brown3" = "GIANA",
#                "darkgoldenrod1" = "clusTCR", 
#                "lightpink" = "GLIPH2",
#                "chartreuse3" = "DeepTCR",
#                "black" = "TCRdist3")

#chain
#m <- m_alpha
m <- m_beta
#m <- m[comb_degree(m) == 9]
cs = comb_size(m)
ss = set_size(m)

color_set <- color_set[which(color_set %in% names(ss))]

#list of methods
list_methods <- list_method_1
#list_methods <- list_method_2
#list_methods <- list_method


UpSet_bis(m,
      comb_col = names(color_set[match(names(ss), color_set)]),
      set_order = list_methods,
      show_column_names = FALSE, lgd_color = color_set,
      right_annotation = rowAnnotation(
      "Set size" = anno_barplot(ss, 
                                ylim = c(0, max(ss)*1.1),
                                border = FALSE, 
                                gp = gpar(fill = names(color_set[match(names(ss), color_set)])), 
                                width = unit(6, "cm"), 
                                add_numbers = TRUE, 
                                numbers_rot = 0, 
                                axis_param = list(side = "top")), 
                                annotation_name_side = "top"),
      top_annotation = HeatmapAnnotation(
      "Intersection size" = anno_barplot(cs, 
                                         ylim = c(0, max(cs)*1.1),
                                         border = FALSE,
                                         gp = gpar(fill = "gray28"),
                                         add_numbers = T, 
                                         numbers_rot = 0, 
                                         height = unit(4, "cm"), 
                                         width = unit(1, "cm"), 
                                         axis_param = list(side = "left")), 
                                         show_legend = TRUE,
                                         annotation_name_side = "left",
                                         annotation_name_rot = 90))



```

# Jaccard Similarity index between sequences clustered by method (not in paper)

```{r Jaccard similarity}
source("../Functions/alignment_functions.R")
chain <-"b"
# 1.sequence similarity between method
list_alignement_HD <- list_alignement_fun("Hamming_Distance", output_HD, chain)
list_alignement_LD <- list_alignement_fun("Levenshtein_Distance", output_LD, chain)
list_alignement_tcrm <- list_alignement_fun("TCRMatch", output_tcrm, chain)
list_alignement_ismart <- list_alignement_fun("iSMART", output_ismart, chain)
list_alignement_giana <- list_alignement_fun("GIANA", output_giana, chain)
list_alignement_clustcr <- list_alignement_fun("clusTCR", output_clustcr_pairing, chain)
list_alignement_gliph <- list_alignement_fun("GLIPH2", output_gliph, chain)
list_alignement_deeptcr <- list_alignement_fun("DeepTCR", output_deeptcr, chain)
list_alignement_tcrdist <- list_alignement_fun("TCRdist", output_tcrdist, chain)

df_HD <- rbindlist(list_alignement_HD) %>% mutate(method = "Hamming_Distance", effectif = 1)
df_LD <- rbindlist(list_alignement_LD) %>% mutate(method = "Levenshtein_Distance", effectif = 1)
df_tcrm <- rbindlist(list_alignement_tcrm) %>% mutate(method = "TCRMatch", effectif = 1)
df_ismart <- rbindlist(list_alignement_ismart) %>% mutate(method = "iSMART", effectif = 1)
df_giana <- rbindlist(list_alignement_giana) %>% mutate(method = "GIANA", effectif = 1)
df_clustcr <- rbindlist(list_alignement_clustcr) %>% mutate(method = "clusTCR", effectif = 1)
df_gliph <- rbindlist(list_alignement_gliph) %>% mutate(method = "GLIPH2", effectif = 1)
df_deeptcr <- rbindlist(list_alignement_deeptcr) %>% mutate(method = "DeepTCR", effectif = 1)
df_tcrdist <- rbindlist(list_alignement_tcrdist) %>% mutate(method = "TCRdist", effectif = 1)

df_cluster_all_1 <- rbind(df_clustcr, df_deeptcr, df_tcrdist, df_gliph)
colnames(df_cluster_all_1)[2] <- "CDR3"
df_cluster_all <- rbind(df_HD, df_LD, df_tcrm, df_ismart, df_giana, df_cluster_all_1)

df_cluster_all <- dcast(df_cluster_all, method ~ CDR3, fun = sum)
df_cluster_all[is.na(df_cluster_all)] <- 0
df_cluster_all <- as.data.frame(df_cluster_all)
df_cluster_all <- remove_first_column(df_cluster_all)

jaccard_matrix <- 1-(as.matrix(vegdist(df_cluster_all, method = "jaccard",binary = T)))

my_palette <- colorRampPalette(brewer.pal(n=7,name= "YlOrRd"))(100)
pheatmap(jaccard_matrix, show_colnames = T, show_rownames = T,  main = "Jaccard similarity index - Beta chain", fontsize_number = 15, display_numbers=T, color = my_palette)


# 2. MDS on Jaccard matrix
mds <- jaccard_matrix %>%
  dist() %>%
  cmdscale() %>%
  as_tibble()
names(mds) <- c("MDS_1", "MDS_2")


# ggscatter(mds, x = "MDS_1", y = "MDS_2", 
#           label = rownames(jaccard_matrix),
#           size = 1,
#           repel = TRUE)

clust <- kmeans(mds, 4)$cluster %>%
  as.factor()
mds <- mds %>%
  mutate(groups = clust)
# Plot and color by groups
p <- ggscatter(mds, x = "MDS_1", y = "MDS_2", 
          label = rownames(jaccard_matrix),
          color = "groups",
          palette = "uchicago",
          size = 2, 
          ellipse = TRUE,
          ellipse.type = "convex",
          repel = TRUE, 
          title = paste("MDS on Jaccard matrix using clustered CDR3", chain), 
          ggtheme = theme_bw())

ggpar(p, legend = "None")
```

# Jaccard Similarity index between clusters (not in paper)

```{r complexeHeatmap jaccard index similarity}
source("../Functions/similarity_functions.R")
# Jaccard between clusters of 2 methods
list_alignement_all <- list("Hamming_Distance" = list_alignement_HD, "Levenshtein_Distance" = list_alignement_LD, "TCRMatch" = list_alignement_tcrm,
                            "iSMART"= list_alignement_ismart, "GIANA" = list_alignement_giana, "clusTCR"= list_alignement_clustcr,
                            "GLIPH2" = list_alignement_gliph, "DeepTCR"= list_alignement_deeptcr, "TCRdist3"= list_alignement_tcrdist)

#list_method_1 <- c("Hamming_Distance","Levenshtein_Distance", "TCRMatch", "iSMART", "GIANA", "clusTCR", "GLIPH2", "DeepTCR", "TCRdist")
#list_method_2 <- c("Hamming_Distance", "Levenshtein_Distance", "TCRMatch", "iSMART", "GIANA", "clusTCR", "GLIPH2", "DeepTCR", "TCRdist")

#example for 1 comparison HD vs LD
list_method_1 <- c("Hamming_Distance")
list_method_2 <- c("Levenshtein_Distance")

my_palette <- colorRampPalette(brewer.pal(n=7,name= "YlOrRd"))(100)

#heatmap of Jaccard matrices
list_cluster_jaccard <- list()
for(method_1 in list_method_1){
  print(paste("method_1", method_1))
  for(method_2 in list_method_2){
    print(paste("method_2", method_2))
    edge_cluster_jaccard <- perform_jaccard_cluster(list_alignement_all[[method_1]], list_alignement_all[[method_2]])
    edge_cluster_jaccard <- edge_cluster_jaccard %>% filter(size_1 >3 & size_2 >3) #filter on big clusters only
    edge_cluster_jaccard <- edge_cluster_jaccard %>% arrange(size_1, size_2)
    
    cluster_jaccard <- edge_cluster_jaccard %>% reshape2::dcast(factor(cluster_1, levels = unique(cluster_1)) ~ factor(cluster_2, levels = unique(cluster_2)), value.var = "jaccard_index") %>% remove_first_column()
    list_cluster_jaccard[[paste(method_1, method_2, sep = "_")]] <- cluster_jaccard

    ann_barplot_cluster_1 <- edge_cluster_jaccard %>% select(cluster_1, size_1) %>% arrange(., size_1) %>% unique()
    rownames(ann_barplot_cluster_1) <- ann_barplot_cluster_1[,1]
    ann_barplot_cluster_1 <- ann_barplot_cluster_1[,2]
    
    ann_barplot_cluster_2 <- edge_cluster_jaccard %>% select(cluster_2, size_2) %>% arrange(., size_2) %>% unique()
    rownames(ann_barplot_cluster_2) <- ann_barplot_cluster_2[,1]
    ann_barplot_cluster_2 <- ann_barplot_cluster_2[,2]
    
    cluster_jaccard <- as.matrix(cluster_jaccard)
    #pdf(paste0(directory, "Plot/alpha_beta_bis/Jaccard_index/Jaccard_", method_1,"_", method_2,"_", chain,"_bigclusters.pdf"), width = 12, height = 10)
    print(Heatmap(cluster_jaccard, name = "Jaccard similarity", column_title = paste(method_2, " clusters - chain", chain),
    row_title = paste(method_1, " clusters"), cluster_rows = F, cluster_columns = F, show_row_names = F, show_column_names = F, col = my_palette,
    right_annotation = rowAnnotation(size = anno_barplot(ann_barplot_cluster_1)), bottom_annotation = columnAnnotation(size=anno_barplot(ann_barplot_cluster_2))))
    #dev.off()
  }
}

```

# Count the number of specificities in clusters for each method (not in paper)

```{r count specificity by cluster, fig.dim=c(12,10)}
# count_spe_by_cluster_HD <- count_spe_by_cluster_one_chain(output_HD) %>%  mutate(method = "Hamming_Distance") %>% as.data.frame()
# count_spe_by_cluster_LD <- count_spe_by_cluster_one_chain(output_LD) %>% mutate(method = "Levenshtein_Distance")%>% as.data.frame()
# count_spe_by_cluster_tcrm <- count_spe_by_cluster_one_chain(output_tcrm) %>% mutate(method = "TCRMatch") %>% as.data.frame()
# count_spe_by_cluster_ismart <- count_spe_by_cluster_one_chain(output_ismart) %>% mutate(method = "iSMART")%>% as.data.frame()
# count_spe_by_cluster_giana <- count_spe_by_cluster_one_chain(output_giana) %>% mutate(method = "GIANA")%>% as.data.frame()
# 
# count_spe_by_cluster_clustcr  <- count_spe_by_cluster_pairing(output_clustcr_pairing, "CDR3b_CDR3a") %>% mutate(method = "clusTCR") %>% as.data.frame()
# count_spe_by_cluster_gliph  <- count_spe_by_cluster_pairing(output_gliph, "CDR3b_CDR3a") %>% mutate(method = "GLIPH2") %>% as.data.frame()
# count_spe_by_cluster_deeptcr  <- count_spe_by_cluster_pairing(output_deeptcr, "CDR3b_CDR3a") %>% mutate(method = "DeepTCR") %>% as.data.frame()
# count_spe_by_cluster_tcrdist <- count_spe_by_cluster_pairing(output_tcrdist, "CDR3b_CDR3a") %>% mutate(method = "TCRdist3")%>% as.data.frame()
# 
# list_count_spe_by_cluster <- list("Hamming_Distance"=count_spe_by_cluster_HD, "Levenshtein_Distance"=count_spe_by_cluster_LD , "TCRMatch"=count_spe_by_cluster_tcrm, "iSMART"=count_spe_by_cluster_ismart, "GIANA"=count_spe_by_cluster_giana, "clusTCR"=count_spe_by_cluster_clustcr, "GLIPH2"=count_spe_by_cluster_gliph, "DeepTCR"=count_spe_by_cluster_deeptcr, "TCRdist3"=count_spe_by_cluster_tcrdist)
# 
# 
# list_plot <- list()
# for(method in names(list_count_spe_by_cluster)){
#   print(method)
#   list_plot[[method]] <- ggplot(data = list_count_spe_by_cluster[[method]] , aes(x= nb_spe_cluster, y = perc_cluster))+
#     geom_bar(stat="identity", fill = color_method[[method]])+
#     theme_bw()+
#     ylab("Percentage of clusters")+
#     xlab("Number of specificities")+
#     ggtitle(method)+
#     ylim(0,100)
# 
# }
# 
# do.call("grid.arrange", c(list_plot, ncol = 3)) 

#Refine the previous analysis by selecting one spe for polyspecific CDR3
count_spe_by_cluster_HD <- select_one_spe_by_seq(output_HD) %>%  mutate(method = "Hamming_Distance") %>% as.data.frame()
count_spe_by_cluster_LD <- select_one_spe_by_seq(output_LD) %>%  mutate(method = "Levenshtein_Distance") %>% as.data.frame()
count_spe_by_cluster_tcrm <- select_one_spe_by_seq(output_tcrm) %>% mutate(method = "TCRMatch") %>% as.data.frame()
count_spe_by_cluster_ismart <- select_one_spe_by_seq(output_ismart) %>% mutate(method = "iSMART")%>% as.data.frame()
count_spe_by_cluster_giana <- select_one_spe_by_seq(output_giana) %>% mutate(method = "GIANA")%>% as.data.frame()

count_spe_by_cluster_clustcr  <- select_one_spe_by_seq_bis(output_clustcr_pairing, "CDR3b_CDR3a") %>% mutate(method = "clusTCR") %>% as.data.frame()
count_spe_by_cluster_gliph  <- select_one_spe_by_seq_bis(output_gliph, "CDR3b_CDR3a") %>% mutate(method = "GLIPH2") %>% as.data.frame()
count_spe_by_cluster_deeptcr  <- select_one_spe_by_seq_bis(output_deeptcr, "CDR3b_CDR3a") %>% mutate(method = "DeepTCR") %>% as.data.frame()
count_spe_by_cluster_tcrdist <- select_one_spe_by_seq_bis(output_tcrdist, "CDR3b_CDR3a") %>% mutate(method = "TCRdist")%>% as.data.frame()


list_count_spe_by_cluster <- list("Hamming_Distance"=count_spe_by_cluster_HD, "Levenshtein_Distance"=count_spe_by_cluster_LD , "TCRMatch"=count_spe_by_cluster_tcrm, "iSMART"=count_spe_by_cluster_ismart, "GIANA"=count_spe_by_cluster_giana, "clusTCR"=count_spe_by_cluster_clustcr, "GLIPH2"=count_spe_by_cluster_gliph, "DeepTCR"=count_spe_by_cluster_deeptcr, "TCRdist3"=count_spe_by_cluster_tcrdist)


list_plot <- list()
for(method in names(list_count_spe_by_cluster)){
  #print(method)
  list_plot[[method]] <- ggplot(data = list_count_spe_by_cluster[[method]] , aes(x= nb_spe_cluster_new, y = perc_cluster))+
    geom_bar(stat="identity", fill = color_method[[method]])+
    theme_bw()+
    ylab("Percentage of clusters")+
    xlab("Number of specificities")+
    ggtitle(method)+
    ylim(0,100)+
    theme(plot.title = element_text(hjust=0.5, face = "bold", size = 20),
          text=element_text(size=15),
        axis.text=element_text(size=15))
}


do.call("grid.arrange", c(list_plot, ncol = 3))

```

# selection of specific epitopes

```{r selection of specific epitopes, fig.dim=c(8, 15)}
count_epitope <- data_human %>% select(CDR3_alpha, CDR3_beta, Epitope) %>% unique()#CDR3_alpha, CDR3_beta
count_epitope <- count_epitope %>% group_by(Epitope) %>% mutate(count = n())
epitope_df <- count_epitope %>% select(Epitope, count) %>% unique()

#plot number of TCR according epitopes
epitope_df <- epitope_df %>% arrange(-count)
epitope_df$Epitope <- factor(epitope_df$Epitope, level = epitope_df$Epitope)
epitope_df_subset <- epitope_df %>% filter(count >=10)

#Select the top 6 epitopes: c("GILGFVFTL", "NLVPMVATV", "GLCTLVAML", "FLCMKALLL", "NYNYLYRLF", "LLFGYPVYV")
epitope_df_subset <- epitope_df_subset %>% as.data.frame()
epitope_df_subset  <- epitope_df_subset %>% mutate(color_epi=ifelse(epitope_df_subset$Epitope=="GILGFVFTL","darkgoldenrod1",
                                                                  ifelse(epitope_df_subset$Epitope=="NLVPMVATV","brown2",
                                                                        ifelse(epitope_df_subset$Epitope=="GLCTLVAML", "pink",
                                                                             ifelse(epitope_df_subset$Epitope=="FLCMKALLL", "cyan2",
                                                                                  ifelse(epitope_df_subset$Epitope=="NYNYLYRLF", "darkorange",
                                                                                     ifelse(epitope_df_subset$Epitope=="LLFGYPVYV","darkolivegreen4","grey50")))))))

#pdf("../Plot/6EpitopeSelection.pdf", height = 12, width = 10)
ggplot(data=epitope_df_subset, aes(x= reorder(Epitope, count), y=count)) +
  geom_bar(stat="identity", aes(fill = color_epi))+
  scale_fill_identity() +
  geom_text(aes(label=count), hjust=-0.3, size=5)+
  theme_classic()+
  theme(plot.title = element_text(hjust= 0.5, size = 16),
        text=element_text(size=20),
        axis.text=element_text(size=15))+
  ggtitle("Count of TCRs")+
  ylab("Count")+
  xlab("Epitope")+
  ylim(0,max(epitope_df_subset$count)+10)+
  coord_flip()
#dev.off()

```

# Processing of output to add specificity

```{r create specific output}
output_HD_spe <- merge(output_HD, dataframe_alpha_beta[,c("CDR3", "Epitope", "chain")], by = "CDR3") %>% unique()
output_HD_spe <- output_HD_spe %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_HD_spe <- output_HD_spe %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])
#output_HD_spe <- output_HD_spe[output_HD_spe$size==output_HD_spe$nb_epi_cluster, ]#select pure clusters

output_LD_spe <- merge(output_LD, dataframe_alpha_beta[,c("CDR3", "Epitope", "chain")], by = "CDR3") %>% unique()
output_LD_spe <- output_LD_spe %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_LD_spe <- output_LD_spe %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])
#output_LD_spe <- output_LD_spe[output_LD_spe$size==output_LD_spe$nb_epi_cluster, ]#select pure clusters

output_tcrm_spe <- merge(output_tcrm, dataframe_alpha_beta[,c("CDR3", "Epitope", "chain")], by = "CDR3") %>% unique()
output_tcrm_spe <- output_tcrm_spe %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_tcrm_spe <- output_tcrm_spe %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])
#output_tcrm_spe <- output_tcrm_spe[output_tcrm_spe$size==output_tcrm_spe$nb_epi_cluster, ]#select pure clusters

output_ismart_spe <- merge(output_ismart, dataframe_alpha_beta[,c("CDR3", "Epitope", "chain")], by = "CDR3") %>% unique()
output_ismart_spe <- output_ismart_spe %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_ismart_spe <- output_ismart_spe %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])
#output_ismart_spe <- output_ismart_spe[output_ismart_spe$size==output_ismart_spe$nb_epi_cluster, ]#select pure clusters

output_giana_spe <- merge(output_giana, dataframe_alpha_beta[,c("CDR3", "Epitope", "chain")], by = "CDR3") %>% unique()
output_giana_spe <- output_giana_spe %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_giana_spe <- output_giana_spe %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])
#output_giana_spe <- output_giana_spe[output_giana_spe$size==output_giana_spe$nb_epi_cluster, ]#select pure clusters

#pairing method
output_clustcr_alpha <- output_clustcr_pairing %>% select(CDR3a, clusters, size) %>% rename("CDR3" = "CDR3a") %>% mutate(chain = "alpha")
output_clustcr_alpha$clusters <- paste(output_clustcr_alpha$clusters, "a", sep ="")#rename clusters alpha
output_clustcr_beta <- output_clustcr_pairing %>% select(CDR3b, clusters, size) %>% rename("CDR3" = "CDR3b") %>% mutate(chain = "beta")
output_clustcr_beta$clusters <- paste(output_clustcr_beta$clusters, "b", sep ="")#rename clusters beta
output_clustcr_spe <- rbind(output_clustcr_alpha, output_clustcr_beta) %>% merge(., dataframe_alpha_beta[,c("CDR3", "Epitope")], by = "CDR3") %>% unique() %>%  group_by(clusters) %>% mutate(size = n())
output_clustcr_spe <- output_clustcr_spe %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_clustcr_spe <- output_clustcr_spe %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])
output_clustcr_spe <- output_clustcr_spe %>% filter(size != 1)
#output_clustcr_spe <- output_clustcr_spe[output_clustcr_spe$size==output_clustcr_spe$nb_epi_cluster, ]#select pure clusters

output_gliph_alpha <- output_gliph %>% select(CDR3a, clusters, size) %>% rename("CDR3" = "CDR3a") %>% mutate(chain = "alpha")
output_gliph_alpha$clusters <- paste(output_gliph_alpha$clusters, "a", sep ="")#rename clusters alpha
output_gliph_beta <- output_gliph %>% select(CDR3b, clusters, size) %>% rename("CDR3" = "CDR3b") %>% mutate(chain = "beta")
output_gliph_beta$clusters <- paste(output_gliph_beta$clusters, "b", sep ="")#rename clusters beta
output_gliph_spe <- rbind(output_gliph_alpha, output_gliph_beta) %>% merge(., dataframe_alpha_beta[,c("CDR3", "Epitope")], by = "CDR3") %>% unique() %>% group_by(clusters) %>% mutate(size = n())
output_gliph_spe <- output_gliph_spe %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_gliph_spe <- output_gliph_spe %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])
output_gliph_spe <- output_gliph_spe %>% filter(size != 1)
#output_gliph_spe <- output_gliph_spe[output_gliph_spe$size==output_gliph_spe$nb_epi_cluster, ]#select pure clusters

output_deeptcr_alpha <- output_deeptcr %>% select(CDR3a, clusters) %>% rename("CDR3" = "CDR3a") %>% mutate(chain = "alpha")
output_deeptcr_alpha$clusters <- paste(output_deeptcr_alpha$clusters, "a", sep ="")#rename clusters alpha
output_deeptcr_beta <- output_deeptcr %>% select(CDR3b, clusters) %>% rename("CDR3" = "CDR3b") %>% mutate(chain = "beta")
output_deeptcr_beta$clusters <- paste(output_deeptcr_beta$clusters, "b", sep ="")#rename clusters beta
output_deeptcr_spe <- rbind(output_deeptcr_alpha, output_deeptcr_beta) %>% merge(., dataframe_alpha_beta[,c("CDR3", "Epitope")], by = "CDR3") %>% unique() %>% group_by(clusters) %>% mutate(size = n())
output_deeptcr_spe <- output_deeptcr_spe %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_deeptcr_spe <- output_deeptcr_spe %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])
output_deeptcr_spe <- output_deeptcr_spe %>% filter(size != 1)
#output_deeptcr_spe <- output_deeptcr_spe[output_deeptcr_spe$size==output_deeptcr_spe$nb_epi_cluster, ]#select pure clusters

output_tcrdist_alpha <- output_tcrdist %>% select(CDR3a, clusters) %>% rename("CDR3" = "CDR3a") %>% mutate(chain = "alpha")
output_tcrdist_alpha$clusters <- paste(output_tcrdist_alpha$clusters, "a", sep ="")#rename clusters alpha
output_tcrdist_beta <- output_tcrdist %>% select(CDR3b, clusters) %>% rename("CDR3" = "CDR3b") %>% mutate(chain = "beta")
output_tcrdist_beta$clusters <- paste(output_tcrdist_beta$clusters, "b", sep ="")#rename clusters beta
output_tcrdist_spe <- rbind(output_tcrdist_alpha, output_tcrdist_beta) %>% merge(., dataframe_alpha_beta[,c("CDR3", "Epitope")], by = "CDR3") %>% unique() %>% group_by(clusters) %>% mutate(size = n())
output_tcrdist_spe <- output_tcrdist_spe %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_tcrdist_spe <- output_tcrdist_spe %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])
output_tcrdist_spe <- output_tcrdist_spe %>% filter(size != 1)
#output_tcrdist_spe <- output_tcrdist_spe[output_tcrdist_spe$size==output_tcrdist_spe$nb_epi_cluster, ]#select pure clusters


list_output_spe <- list("Hamming_Distance" = output_HD_spe, "Levenshtein_Distance"=output_LD_spe, "TCRMatch"= output_tcrm_spe, "GIANA"= output_giana_spe, "iSMART" = output_ismart_spe, "clusTCR" = output_clustcr_spe, "GLIPH2" = output_gliph_spe,"DeepTCR" = output_deeptcr_spe, "TCRdist3" = output_tcrdist_spe)

```

# How many epitope-specific sequences are clustered by methods ?

```{r How many sequences/pairs are clustered by methods }
output_clustcr_spe_bis <- output_clustcr_pairing %>% merge(., data_human[,c("CDR3b_CDR3a", "Epitope")], by.x = "CDR3b_CDR3a", by.y="CDR3b_CDR3a") %>% unique()
colnames(output_clustcr_spe_bis)[1] <- "CDR3b_CDR3a"

output_gliph_spe_bis <- output_gliph %>% merge(., data_human[,c("CDR3b_CDR3a", "Epitope")], by="CDR3b_CDR3a") %>% unique()

output_deeptcr_spe_bis <- output_deeptcr %>% merge(., data_human[,c("CDR3b_CDR3a", "Epitope")], by="CDR3b_CDR3a") %>% unique()

output_tcrdist_spe_bis <- output_tcrdist %>% merge(., data_human[,c("CDR3b_CDR3a", "Epitope")], by="CDR3b_CDR3a") %>% unique() 


list_output_spe_bis <- list("Hamming_Distance" = output_HD_spe, "Levenshtein_Distance"=output_LD_spe, "TCRMatch"= output_tcrm_spe, "GIANA"= output_giana_spe, "iSMART" = output_ismart_spe, "clusTCR" = output_clustcr_spe_bis, "GLIPH2" = output_gliph_spe_bis,"DeepTCR" = output_deeptcr_spe_bis, "TCRdist3" = output_tcrdist_spe_bis)


count_spe_method <- data.frame(epitope = c("GILGFVFTL", "NLVPMVATV", "GLCTLVAML", "FLCMKALLL", "NYNYLYRLF", "LLFGYPVYV"),
                               total_alpha= c(484, 325,145,134,55,4),
                               total_beta=c(456,349,170,136,46,7),
                               total_pair = c(706,397,215,136,56,10))
count_spe_method <- count_spe_method %>% mutate(total_alpha_plus_beta = total_alpha + total_beta)

count_spe_method_epi <- data.frame()
for(method in list_method){
  #print(method)
  big_df <- data.frame()
  epitope_list <- list()
  for(epi in list_epitope){
    #print(epi)
    test <- list_output_spe_bis[[method]] %>% filter(Epitope == epi)
    if(method %in% c("Hamming_Distance","Levenshtein_Distance","TCRMatch","iSMART","GIANA")){
      nb_cdr3 <- n_distinct(test$CDR3)
      df <- data.frame(methods = method, epitope = epi, nb_clustered = nb_cdr3, perc_clustered = nb_cdr3/count_spe_method$total_alpha_plus_beta[count_spe_method$epitope == epi]*100)
    }
    else{
      nb_cdr3 <- n_distinct(test$CDR3b_CDR3a)
      df <- data.frame(methods = method, epitope = epi, nb_clustered = nb_cdr3, perc_clustered = nb_cdr3/count_spe_method$total_pair[count_spe_method$epitope == epi]*100)
    }
    
    count_spe_method_epi <- rbind(count_spe_method_epi , df)
  }
}

count_spe_method_epi$methods <- factor(count_spe_method_epi$methods, level = list_method)
count_spe_method_epi$epitope <- factor(count_spe_method_epi$epitope, level = list_epitope)

# ggplot(data = count_spe_method_epi, aes(x= methods, y = perc_clustered, fill = epitope))+
#   geom_bar(stat="identity", position=position_dodge())+
#   theme_bw()+
#   scale_fill_manual(values =c("darkgoldenrod1", "brown2", "pink", "cyan2", "darkorange", "darkolivegreen4"))+
#   ylab("Percentage of clustered seq/pair")+
#   xlab("Methods")+
#   theme(legend.position = "bottom")


list_plot <- list()
for(epi in list_epitope){
  df <- count_spe_method_epi %>% filter(epitope ==epi)
  df$methods <- factor(df$methods, levels = list_method)
  
  list_plot[[epi]] <- ggplot(data = df, aes(x= methods, y = perc_clustered, fill = methods))+
    geom_bar(stat="identity", position=position_dodge())+
    theme_bw()+
    scale_fill_manual(values = color_method)+
    ylab("Percentage of clustered seq/pair (%)")+
    xlab("Methods")+
    theme(legend.position = "none",
          #axis.title.x = element_blank(),axis.text.x = element_blank())+
          axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(hjust=0.5, face = "bold", size =20),
          text=element_text(size=15),
        axis.text=element_text(size=13))+
    ggtitle(epi)+xlab("")
}
#do.call("grid.arrange", c(list_plot, ncol = 3))
#pdf("../Plot/PercClusteredSeq_GIL.pdf", height= 6, width = 7)
grid.arrange(list_plot[["GILGFVFTL"]])
#dev.off()
#grid.arrange(list_plot[["FLCMKALLL"]])
```

# Selection of specific clusters

```{r specific clusters of random 6 epitopes}
list_epitope <- c("GILGFVFTL", "NLVPMVATV", "GLCTLVAML", "FLCMKALLL", "NYNYLYRLF", "LLFGYPVYV")

#1. with at least 1 specificity
at_least_one_epi <- data.frame()
at_least_one_epi_clusters <-list()
for(method in list_method){
  #print(method)
  big_df <- data.frame()
  epitope_list <- list()
  for(epi in list_epitope){
    #print(epi)
    test <- list_output_spe[[method]] %>% filter(Epitope == epi)#Attention: que sequence spe de GIL (pas les autres du cluster)
    list_clusters <- unique(test$clusters)#retrieve the list of clusters
      
    cluster_spe <- list_output_spe[[method]] %>% filter(clusters %in%list_clusters)
    
    #cluster_spe <- cluster_spe %>% filter(Epitope != epi) %>% unique()#remove specific CDR3 but for poly CDR3, remove only the row with the specificity not the other one
    #list_CDR3_to_remove <- cluster_spe$CDR3[cluster_spe$Epitope == epi]
    #cluster_spe <- cluster_spe %>% filter(CDR3 %!in% list_CDR3_to_remove)#remove all CDR3 which are specific to the epi
               
    #cluster_spe <- cluster_spe %>% group_by(clusters) %>% mutate(size = n_distinct(CDR3))#resize clusters
    
    epitope_list[[epi]] <- cluster_spe %>% mutate(Epitope_cluster = epi)
    percentage_clusters <- n_distinct(test$clusters)/n_distinct(list_output_spe[[method]]$clusters)
    df <- data.frame(method = method, epitope = epi, perc_clusters = percentage_clusters*100) 
    big_df <- rbind(big_df, df)
  }
  at_least_one_epi <- rbind(at_least_one_epi, big_df)
  at_least_one_epi_clusters[[method]] <- epitope_list
}


at_least_one_epi$method <- factor(at_least_one_epi$method, level = list_method)
at_least_one_epi$epitope <- factor(at_least_one_epi$epitope, level = list_epitope)

# ggplot(data = at_least_one_epi , aes(x= method, y = perc_clusters, fill = epitope))+
#   geom_bar(stat="identity", position=position_dodge())+
#   theme_bw()+
#   scale_fill_manual(values =c("darkgoldenrod1", "brown2", "pink", "cyan2", "darkorange", "darkolivegreen4"))+
#   ylab("Percentage of clusters")+
#   xlab("Methods")+
#   theme(legend.position = "bottom")+
#   ylim(0,50)


# list_plot <- list()
# for(epi in list_epitope){
#   df <- at_least_one_epi %>% filter(epitope ==epi)
#   list_plot[[epi]] <- ggplot(data = df, aes(x= method, y = perc_clusters, fill = method))+
#     geom_bar(stat="identity", position=position_dodge())+
#     theme_bw()+
#     scale_fill_manual(values = color_method)+
#     ylab("Percentage of clusters")+
#     xlab("Methods")+
#     theme(legend.position = "none",
#           axis.title.x = element_blank(),axis.text.x = element_blank())+
#     ggtitle(epi)+ylim(0,50)
# }
# do.call("grid.arrange", c(list_plot, ncol = 3)) 


# for(method in names(at_least_one_epi_clusters)){
#     print(method)
#     df_all_epi <- bind_rows(at_least_one_epi_clusters[[method]])
#     df_all_epi_2 <- df_all_epi %>% select(-c(CDR3, nb_epi_cluster, Epitope, epi_majoritaire)) %>% unique()
#     df_all_epi_2 <- df_all_epi_2 %>% mutate(interval_size = ifelse(size<4, "[2,3]", 
#                                                      ifelse(size >3 & size<11, "[4-10]",
#                                                           ifelse(size>10 & size<51, "[11-50]",
#                                                                  ifelse(size>50 & size <101, "[50,100]",
#                                                                         ifelse(size>100, ">100", NA))))))
#     
#     df_all_epi_2 <- df_all_epi_2 %>% group_by(interval_size, Epitope_cluster) %>% mutate(count_cluster = n()) 
#     df_all_epi_2 <- df_all_epi_2 %>% group_by(interval_size) %>% mutate(count_cluster_in_interval = n()) 
#     df_all_epi_2 <- df_all_epi_2 %>% mutate(perc_spe_per_inter_size = count_cluster/count_cluster_in_interval*100)
#     df_all_epi_3 <- df_all_epi_2 %>% select(Epitope_cluster, interval_size, count_cluster, perc_spe_per_inter_size, count_cluster_in_interval) %>% unique() %>% mutate(method = method)
#     
#     df_all_epi_3$interval_size <- factor(df_all_epi_3$interval_size, level = c("[2,3]", "[4-10]", "[11-50]",  "[50,100]", ">100"))
#     df_all_epi_3$Epitope_cluster <- factor(df_all_epi_3$Epitope_cluster, level = c("GILGFVFTL", "NLVPMVATV", "GLCTLVAML", "FLCMKALLL", "NYNYLYRLF", "LLFGYPVYV"))
#     
#     
#     #pdf(paste0(directory, "Plot/alpha_beta_bis/perc_clusters_one_perc_by_size_", method, "_bis.pdf"), width = 10, height = 6)
#     print(ggplot(data = df_all_epi_3, aes(x=interval_size, y = perc_spe_per_inter_size, fill = Epitope_cluster))+
#     geom_bar(stat="identity", alpha = 1)+
#     scale_fill_manual(values = c("darkgoldenrod1", "brown2", "pink", "cyan2", "darkorange", "darkolivegreen4"))+
#     geom_text(data = df_all_epi_3, aes(x = interval_size, y = 105, label = count_cluster_in_interval), size = 5)+
#     theme_classic()+
#     ggtitle(paste0("Specific clusters (at least 1 ) of ", method)))
#     #dev.off()
# }


# Refine clusters with at least one spe by selecting one spe by sequence
at_least_one_epi_clusters_bis <- list()
for(method in names(at_least_one_epi_clusters)){
  #print(method)
  at_least_one_epitope <- list()
  for(epi in list_epitope){
    output_spe_test <- split(at_least_one_epi_clusters[[method]][[epi]], at_least_one_epi_clusters[[method]][[epi]]$clusters)
    output_spe_new <- list()
    for(cluster in names(output_spe_test)){
      output_spe_seq <- split(output_spe_test[[cluster]], output_spe_test[[cluster]]$CDR3)
      row_toKeep_list <- list()
      for(i in names(output_spe_seq)){
        #print(i)
        if(nrow(output_spe_seq[[i]])>1){#if multiple specificities
          if(output_spe_seq[[i]]$Epitope==output_spe_seq[[i]]$epi_majoritaire[1]){
            row_toKeep_df <- output_spe_seq[[i]] %>% filter(Epitope == output_spe_seq[[i]]$epi_majoritaire[1])#keep specific epitope
          }else{row_toKeep_df <- output_spe_seq[[i]][1,]}#otherwise keep the first row
          
        }else{row_toKeep_df <- output_spe_seq[[i]][1,]}#otherwise keep the first row
        
        row_toKeep_list[[i]] <- row_toKeep_df 
      }
      
      output_spe_keep <- rbindlist(row_toKeep_list) %>% unique() %>% drop_na() %>% as.data.frame()
      output_spe_new[[cluster]] <- output_spe_keep
    }
    
    output_spe_bis <- rbindlist(output_spe_new)
    
    at_least_one_epitope[[epi]] <- output_spe_bis
  }
  
  at_least_one_epi_clusters_bis[[method]] <- at_least_one_epitope
}

```

# Look at the other specificities in specific clusters

```{r other specificities in spe clusters, fig.dim=c(15,10)}
epi <- "GILGFVFTL"
df_organism_color <- data.frame(Antigen_organism = c("SARS-CoV2","Cytomegalovirus (CMV)","Epstein Barr virus (EBV)", "Influenza A virus", "Mycobacterium tuberculosis" ,
                            "Yellow fever virus (YFV)","Homo sapiens (human)","SARS-CoV1","Hepatitis B virus (HBV)","SARS coronavirus Tor2","Tumor associated antigen (TAA)"),
                            color = c("#FFFF99","#7FC97F","#BEAED4", "#FDC086","#F0027F","#386CB0","#666666","darkcyan", "#BF5B17", "aquamarine2","grey"))
df_organism <- merge(df_organism_color, epitope_organism[,c("Antigen_organism")], by = "Antigen_organism", all = T) %>% unique()
df_organism$color[is.na(df_organism$color)] <- "black"
df_organism$Antigen_organism[df_organism$Antigen_organism=="Influenza A virus (A/chicken/Anhui/1/1998(H9N2)) (Influenza A virus (A/Chicken/Anhui/1/98(H9N2)))"]<- "Influenza A virus (chicken)"

df_all <- list()
list_plot <- list()
df_count_spe <- data.frame()
for(method in names(at_least_one_epi_clusters_bis)){
  #print(method)
  df <- at_least_one_epi_clusters_bis[[method]][[epi]] 
  
  if(nrow(df)!=0){
  df <- df %>% select(CDR3, Epitope) %>% group_by(Epitope) %>% mutate(count = n())%>% mutate(method = method)
  df <- df %>% select(Epitope, count, method) %>% unique()
  
  df_spe <- data.frame(epitope = epi, method = method, n_distinct_spe = n_distinct(df$Epitope))
  df_count_spe <- rbind(df_count_spe, df_spe)
  
  df <- df[order(df$count, decreasing = T),]
  if(nrow(df)>10){  df <- df[1:10,]}#show uniqly the 10 first rows
  
  df <- merge(df, epitope_organism, by ="Epitope") %>% unique()
  df <- merge(df, df_organism, by = "Antigen_organism")
  df_color <- df %>% select(Antigen_organism, color) %>% unique()
  #df_color$Antigen_organism[is.na(df_color$Antigen_organism)] <- "grey"
  
  df_all[[method]] <- df
  
  nb_max <- max(df_all[[method]]$count)

  if(nb_max >1000){th <- 200}else if(nb_max >100){th <- 30} else{th <- 1}
  
  list_plot[[method]] <- ggplot(data=df_all[[method]], aes(x= reorder(Epitope, count), y=count, fill = Antigen_organism)) +
  geom_bar(stat="identity")+
  #scale_fill_identity() +
  scale_fill_manual(values = df_color$color)+
  geom_text(aes(label=count), hjust=-0.3, size=4)+
  theme_classic()+
  theme(plot.title = element_text(hjust= 0.5, size = 20, face ="bold"),
        plot.margin=unit(c(0.5,0.5,0.5,0.5),"cm"),
        #legend.position="right")+
        legend.position = c(1, 0.05),
        legend.justification = c("right", "bottom"),
        legend.box.just = "right",
        legend.margin = margin(0.5, 0.5, 0.5, 0.5),
        legend.key.size = unit(0.3, 'cm'), #change legend key size
        legend.key.height = unit(0.3, 'cm'), #change legend key height
        legend.key.width = unit(0.3, 'cm'),
        text=element_text(size=15),
        axis.text=element_text(size=13),
        legend.text=element_text(size=10),
        legend.title=element_text(size=10))+
  ggtitle(method)+
  ylab("Count")+
  xlab("Epitope")+
  ylim(0, max(df_all[[method]]$count)+th)+
  #annotate("text", x=6, y=max(df_all[[method]]$count), label= n_distinct(df_all[[method]]$Epitope))+
  coord_flip()#7x9
  }
}
#pdf("../Plot/AllSpecificitiesInSpeClusters_GIL.pdf", height = 10, width= 15)
do.call("grid.arrange", c(list_plot, ncol = 3))
#dev.off()

```

# Look at in how many clusters selected CDR3s are clustered

```{r number of clusters for selected epitope}
#for selected CDR3, In how many clusters are CDR3 clustered ? 
plot_list <- list()
for(epi in list_epitope){
  df_count_clusters <- data.frame()
  for(method in names(at_least_one_epi_clusters)){
    #print(method)
    df <- at_least_one_epi_clusters[[method]][[epi]] %>% filter(Epitope == epi) %>% select(CDR3, clusters)%>% mutate(method = method) %>% unique()
    
    df_cluster <- data.frame(epitope = epi, method = method, n_distinct_clusters = n_distinct(df$clusters))
    df_count_clusters <- rbind(df_count_clusters, df_cluster)
  }
  df_count_clusters$epitope <- factor(df_count_clusters$epitope, level = list_epitope)
  df_count_clusters$method <- factor(df_count_clusters$method, level = list_method)
  
  plot_list[[epi]] <- ggplot(data= df_count_clusters, aes(x=method, y=n_distinct_clusters, fill = method))+
                  geom_bar(stat="identity")+
                  scale_fill_manual(values = color_method)+
                  theme_bw()+
                  ggtitle(epi)+
                  ylab("cluster count")+
                  xlab(" ")+
                  theme(legend.position ="none",
                        axis.text.x = element_blank())
}
#do.call("grid.arrange", c(plot_list, ncol = 3)) 

#add the notion of cluster size
cluster_total_list <- c("Hamming_Distance"=548, "Levenshtein_Distance"=494, "TCRMatch"=285, "iSMART"=465,"GIANA"=593, "clusTCR"=111*2,"GLIPH2"=380*2, "DeepTCR"=1628*2, "TCRdist3"=669*2)

plot_list <- list()
for(epi in list_epitope){
  df_count_clusters <- data.frame()
  for(method in names(at_least_one_epi_clusters)){
    #print(method)
    df <- at_least_one_epi_clusters[[method]][[epi]] %>% filter(Epitope == epi) %>% select(clusters, size)%>% mutate(method = method) %>% unique()#with spe CDR3
    if(nrow(df)!=0){
    
    df_cluster <- df %>% mutate(interval_size = ifelse(size<4, "[2-3]", 
                                                     ifelse(size >3 & size<11, "[4-10]",
                                                          ifelse(size>10 & size<51, "[11-50]",
                                                                 ifelse(size>50 & size <101, "[50-100]",
                                                                        ifelse(size>100, ">100", NA))))))
    df_cluster <- df_cluster %>% group_by(interval_size) %>% mutate(count = n(), epitope = epi)
    #with percentage of clusters instead of count
    nb <- cluster_total_list[[method]]
    #df_cluster <- df_cluster %>% mutate(nb_total_cluster = nb) #for percentage in the total count of clusters by method
    df_cluster <- df_cluster %>% mutate(nb_total_cluster = n_distinct(df_cluster$clusters)) #for percentage in the count of specific clusters
    df_cluster <- df_cluster %>% mutate(perc_cluster = count/nb_total_cluster*100)
    #
    df_cluster <- df_cluster %>% select(-c(clusters, size)) %>% unique()
      
    df_count_clusters <- rbind(df_count_clusters, df_cluster)
    }
  }
  df_count_clusters$epitope <- factor(df_count_clusters$epitope, level = list_epitope)
  df_count_clusters$method <- factor(df_count_clusters$method, level = list_method)
  df_count_clusters$interval_size <- factor(df_count_clusters$interval_size, levels= rev(c("[2-3]", "[4-10]", "[11-50]", "[50-100]",">100")))
  
  plot_list[[epi]] <- ggplot(data = df_count_clusters, aes(x=method, y = perc_cluster, fill = interval_size))+#y=count
  geom_bar(stat="identity")+
  scale_fill_grey(start=0.8, end=0.2) +
  theme_bw()+
    ggtitle(epi)+
    ylab("Percentage of clusters (%)")+
    xlab(" ")+
    theme(legend.position ="right",
          #axis.text.x = element_blank())+
          axis.text.x = element_text(hjust=1, angle = 45),
          plot.title = element_text(hjust=0.5, face = "bold", size = 20),
          text=element_text(size=15),
        axis.text=element_text(size=13))+
    ylim(0,100)

}

#do.call("grid.arrange", c(plot_list, ncol = 3))
#pdf("../Plot/PercClusters_GIL.pdf", height = 6, width = 8)
grid.arrange(plot_list[["GILGFVFTL"]])
#dev.off()
#grid.arrange(plot_list[["FLCMKALLL"]])
```

# Number of specificities in specific clusters

```{r Number of specificities in specific clusters, fig.dim=c(12,10)}
#Number of specificities in specific clusters by selecting one spe for multispe CDR3
nb_spe_cluster_spe_method <-list()
for(method in list_method){
  #print(method)
  nb_spe_cluster_spe <- list()
  for(epi in list_epitope){
    #print(epi)
    if(nrow(at_least_one_epi_clusters_bis[[method]][[epi]] != 0)){
      at_least_one_epi_clusters_bis[[method]][[epi]] <- at_least_one_epi_clusters_bis[[method]][[epi]] %>% group_by(clusters) %>% mutate(nb_spe_cluster = n_distinct(Epitope))
      cluster_spe <- at_least_one_epi_clusters_bis[[method]][[epi]] %>% select(clusters, Epitope, nb_spe_cluster) %>% unique()
      cluster_spe <- cluster_spe %>% select(-Epitope) %>% unique()
      
      df_sum <- cluster_spe %>% group_by(nb_spe_cluster) %>% summarise(count_cluster = n(), .groups = 'drop')
      cluster_total <- sum(df_sum$count_cluster)
      df_sum <- df_sum %>% mutate(perc_cluster = count_cluster/cluster_total*100)
      
      df_sum <- df_sum %>% mutate(methods = method, epitope_spe = epi, cluster_tot=cluster_total)
      
      nb_spe_cluster_spe[[epi]] <- df_sum
    }
  }
  nb_spe_cluster_spe_method[[method]] <- nb_spe_cluster_spe
}

list_color_epitope <- c("darkgoldenrod1", "brown2", "pink", "cyan2", "darkorange", "darkolivegreen4")
epi <- "GILGFVFTL"
n <- 1
list_plot <- list()
for(method in list_method){
  df <- nb_spe_cluster_spe_method[[method]][[epi]]
  if(exists("df")&&is.data.frame(get("df"))==TRUE){
  
  list_plot[[method]] <- ggplot(data = df, aes(x= nb_spe_cluster, y = perc_cluster, fill = epitope_spe))+
        geom_bar(stat="identity")+
        theme_bw()+
        ylab("Percentage of clusters")+
        xlab("Number of specificities")+
        scale_fill_manual(values = list_color_epitope[[n]])+
        ggtitle(method)+
        ylim(0,100)+
      theme(strip.background =element_rect(fill="grey"),legend.position="none", 
            plot.title = element_text(hjust = 0.5, face = "bold", size = 20),
            text=element_text(size=15),
        axis.text=element_text(size=15))
  }
}
#pdf("../Plot/NumberSpecificities_GIL.pdf", height = 10, width = 15)
do.call("grid.arrange", c(list_plot, ncol = 3))
#dev.off()
```

# Selection of very specific clusters and calculation of sensibility

```{r selection of specific clusters - major epitope and x times more - sensibility}
output_clustcr_spe_ter <- output_clustcr_spe_bis %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_clustcr_spe_ter <- output_clustcr_spe_ter %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])

output_gliph_spe_ter <- output_gliph_spe_bis %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_gliph_spe_ter <- output_gliph_spe_ter %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])

output_deeptcr_spe_ter <- output_deeptcr_spe_bis %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_deeptcr_spe_ter <- output_deeptcr_spe_ter %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])

output_tcrdist_spe_ter <- output_tcrdist_spe_bis %>% group_by(clusters, Epitope) %>% mutate(nb_epi_cluster = n())
output_tcrdist_spe_ter <- output_tcrdist_spe_ter %>% group_by(clusters)  %>% mutate(epi_majoritaire = Epitope[which.max(nb_epi_cluster)])

list_output_spe_ter <- list("Hamming_Distance" = output_HD_spe, "Levenshtein_Distance"=output_LD_spe, "TCRMatch"= output_tcrm_spe, "GIANA"= output_giana_spe, "iSMART" = output_ismart_spe, "clusTCR" = output_clustcr_spe_ter, "GLIPH2" = output_gliph_spe_ter,"DeepTCR" = output_deeptcr_spe_ter, "TCRdist3" = output_tcrdist_spe_ter)

#1. select clusters spe
major_times_epi_clusters <-list()
for(method in list_method){
  #print(method)
  big_df <- data.frame()
  epitope_list <- list()
  for(epi in list_epitope){
    #print(epi)
    list_output_spe_ter[[method]] <- list_output_spe_ter[[method]] %>% mutate(perc_epi_in_cluster = nb_epi_cluster/size)
    test <- list_output_spe_ter[[method]] %>% filter(epi_majoritaire == epi & size >3 & perc_epi_in_cluster >=0.67)#select clusters with epi as major and bigcluster
    list_clusters <- unique(test$clusters)#retrieve the list of clusters
      
    cluster_spe <- list_output_spe_ter[[method]] %>% filter(clusters %in% list_clusters)
    if(method =="GLIPH2"){cluster_spe  <- cluster_spe %>% filter(` number_unique_cdr3`>3)}
    cluster_spe$CDR3b_CDR3a_clusters <- paste(cluster_spe$CDR3b_CDR3a, cluster_spe$clusters, sep = "_")
    
    epitope_list[[epi]] <- cluster_spe %>% mutate(major_xtimes_epi = epi)

  }

  major_times_epi_clusters[[method]] <- epitope_list
}

#2. calculation of the sensitivity
df_nb_spe <- data.frame(epitope = list_epitope,
                       nb_CDR3 =c(sum(matrix_alpha_beta_final$GILGFVFTL), sum(matrix_alpha_beta_final$NLVPMVATV), sum(matrix_alpha_beta_final$GLCTLVAML), 
                                  sum(matrix_alpha_beta_final$FLCMKALLL), sum(matrix_alpha_beta_final$NYNYLYRLF), sum(matrix_alpha_beta_final$LLFGYPVYV)),
                       nb_pair =c(sum(matrix_pair_final$GILGFVFTL), sum(matrix_pair_final$NLVPMVATV), sum(matrix_pair_final$GLCTLVAML), 
                                  sum(matrix_pair_final$FLCMKALLL), sum(matrix_pair_final$NYNYLYRLF), sum(matrix_pair_final$LLFGYPVYV)))

df_sensibility_all_epitope <- data.frame()
for (epi in list_epitope){
  df_sensibility_all <- data.frame()
  for(method in list_method){
    if(method %in% c("Hamming_Distance", "Levenshtein_Distance", "TCRMatch","iSMART", "GIANA")){
      nb_CDR3 <- n_distinct(major_times_epi_clusters[[method]][[epi]]$CDR3[major_times_epi_clusters[[method]][[epi]]$Epitope==epi])
      score <- nb_CDR3/df_nb_spe$nb_CDR3[df_nb_spe$epitope==epi]
    }else{
      nb_pair <- n_distinct(major_times_epi_clusters[[method]][[epi]]$CDR3b_CDR3a[major_times_epi_clusters[[method]][[epi]]$Epitope==epi])
      score <- nb_pair/df_nb_spe$nb_pair[df_nb_spe$epitope==epi]
    }
      df_sensibility <- data.frame(method = method, epitope = epi, sensibility = score)
      df_sensibility_all <- rbind(df_sensibility_all, df_sensibility)
  }
  df_sensibility_all_epitope <- rbind(df_sensibility_all_epitope, df_sensibility_all)
}

#by epitope
df_sensibility_all_epitope$method <- factor(df_sensibility_all_epitope$method, levels = list_method)
df_sensibility_all_epitope$epitope <- factor(df_sensibility_all_epitope$epitope, levels = list_epitope)

# ggplot(df_sensibility_all_epitope, aes(x=method, y=sensibility, color = method)) +
#   geom_segment(aes(xend=method, yend=0)) +
#   geom_point(size=4) +
#   scale_color_manual(values = color_method)+
#   theme_bw() +
#   xlab("")+
#   theme(axis.text.x =  element_blank())+
#   facet_wrap(~epitope)

#for GIL epitope only
#pdf("../Plot/Sensibility_GIL.pdf", height=6, width=7)
ggplot(df_sensibility_all_epitope[df_sensibility_all_epitope$epitope == "GILGFVFTL",], aes(x=method, y=sensibility,fill = method)) +
  # geom_segment(aes(xend=method, yend=0)) +
  # geom_point(size=4) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  scale_color_manual(values = color_method)+
  theme_bw() +
  scale_fill_manual(values = c("darkorange", "cyan3", "darkorchid4", "grey", "brown1", "darkgoldenrod1", "pink", "chartreuse3", "black"))+
  geom_text(aes(label = round(sensibility, 2)), vjust = -0.5,
             position = position_dodge(0.9)) +
  xlab("")+
  theme(axis.text.x =  element_text(angle =45, hjust = 1),
        legend.position="none",
        plot.title = element_text(hjust = 0.5, face = "bold", size = 20),
        text=element_text(size=15),
        axis.text=element_text(size=15))+
  ggtitle("GILGFVFTL")+ylim(0,1)+ylab("sensitivity")
#dev.off()

#boxplot
library(rstatix)
stat.test <- df_sensibility_all_epitope %>% wilcox_test(sensibility ~ method)

boxplot <- ggplot(data= df_sensibility_all_epitope, aes(x=method, y=sensibility))+
    geom_boxplot()+
    geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, aes(fill = epitope))+
    theme_bw()+
    ylim(0,1)+
    theme(axis.text.x = element_text(angle =45, hjust = 1),
          legend.position="right",
          legend.text=element_text(size=15),
          legend.title=element_text(size=15),
          text=element_text(size=15),
          axis.text=element_text(size=15))+
    scale_fill_manual(values= list_color_epitope)+xlab("")+ylab("sensitivity")#5x5

stat.test <- stat.test %>% add_xy_position(x = "method")
stat.test$custom.label <- ifelse(stat.test$p.adj <= 0.05, stat.test$p.adj, "ns")

boxplot <- boxplot + stat_pvalue_manual(stat.test, label = "custom.label", tip.length = 0.01)#non significatif

# pdf("../Plot/SensibilityAllEpitopes.pdf", height=6, width =8)
# boxplot
# dev.off()

```

# R Session Info

```{r}
utils::sessionInfo()
```

